---
title: Function Test Discrete Allocation with Linear Regression
description: |
  Test binary optimal allocation solution function from package.
  Use the California student test score dataset, same as line by line.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function Test Discrete Allocation with Linear Regression}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---


Test discrete optimal allocation solution line by line without function. Use the California student test score dataset. Regress student English and Math test scores on Student-Teacher-Ratio.


This function produces the same results as [DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Line by Line)](https://fanwangecon.github.io/PrjOptiAlloc/articles/ffv_opt_sodis_rkone_casch_allrw.html). The differences are:

1. the program here uses the dataframes created by the line by line file as Inputs
2. the program here uses functions that are identical to the line by line function, but are saved in the R folder.

This file, by invoking functions, is much shorter than the line by line program.


```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)

library(PrjOptiAlloc)

library(knitr)
library(kableExtra)
```

## Load Pre-Generated Data

These are data thata were generated in the [DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Line by Line)](https://fanwangecon.github.io/PrjOptiAlloc/articles/ffv_opt_sodis_rkone_casch_allrw.html) function.

But here, select only a small subset of data to clarity and debugging. Select on 3 school districts. Select them randomly. 

First, load data:

```{r}
# Load Data
data(df_opt_caschool_prep_i)
data(df_opt_caschool_input_il)
data(df_opt_caschool_input_ib)

# Rescaling to avoid out of possibility calculations
fl_rescale_ratio = 0.01
df_opt_caschool_input_il <- df_opt_caschool_input_il %>%
  mutate(A_il = A_il*fl_rescale_ratio, 
         alpha_il = alpha_il*fl_rescale_ratio)
df_opt_caschool_input_ib <- df_opt_caschool_input_ib %>%
  mutate(A_i_l0 = A_i_l0*fl_rescale_ratio, 
         alpha_o_i = alpha_o_i*fl_rescale_ratio)
```

Second, select a subset of rows form dataframe with i-specific rows: 

```{r}
# Select three random school IDs
set.seed(11)
it_select <- dim(df_opt_caschool_prep_i)[1]
ar_it_sample_rows <- sample(dim(df_opt_caschool_prep_i)[1],
                            it_select, replace=FALSE)
df_opt_caschool_prep_i <- df_opt_caschool_prep_i[ar_it_sample_rows,]
ar_selected_keys <- sort(df_opt_caschool_prep_i %>% pull(id_i))
print(ar_selected_keys)
```

Third select only based on subset of keys from the three dataframes:

```{r}
df_opt_caschool_prep_i <- df_opt_caschool_prep_i %>% filter(id_i %in% ar_selected_keys)
df_opt_caschool_input_il <- df_opt_caschool_input_il %>% filter(id_i %in% ar_selected_keys)
df_opt_caschool_input_ib <- df_opt_caschool_input_ib %>% filter(id_i %in% ar_selected_keys)
```

## Solve All Districts Discrete Optimal Allocation Across Planner Preferences

### Specify Policy Parameters

Note that *input_il* function already is based on *fl_fi_max*.

```{r control policy}
# 100 percent teacher at most per school☺, discretize floor as needed
# This is not○ a parameter t•hat matters here, already a part of the input_il function
fl_fi_max = 0.20
# 20 percent total additional of all teachers
fl_fa_max = 0.05
# What is the number of teachers we can increase by
fl_teacher_increase_number <- sum(df_opt_caschool_prep_i$teachers)*fl_fa_max
fl_teacher_increase_number <- floor(fl_teacher_increase_number)

# Rho values to consider
ar_rho <- 1 - (10^(c(seq(-2.2,2.2, length.out=60))))
ar_rho <- unique(ar_rho)
```

## Optimal Queue and Allocations

Use the discrete allocation function across preference [ffp_opt_anlyz_rhgin_dis.html](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin_dis.html) function. This computes optimal allocation for multiple planner inequality aversion $\lambda$ parameters at the same time. Note that in the function $\lambda=\rho$.

```{r}
ls_df_queue <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho,
                                       fl_teacher_increase_number,
                                       bl_df_alloc_il = TRUE,
                                       df_input_il = df_opt_caschool_input_il,
                                       svr_rho = 'rho',
                                       svr_id_i = 'id_i', svr_id_il = 'id_il',
                                       svr_D_max_i = 'D_max_i', svr_D_il = 'D_il',
                                       svr_D_star_i = 'D_star_i', svr_F_star_i = 'F_star_i',
                                       svr_inpalc = 'Q_il', svr_D_Wbin_il = 'D_Wbin_il',
                                       svr_A_il = 'A_il', svr_alpha_il = 'alpha_il', svr_beta_i = 'beta_i',
                                       svr_expout = 'opti_exp_outcome',
                                       st_idcol_prefix = 'sid_')
df_queue_il_long <- ls_df_queue$df_queue_il_long
df_queue_il_wide <- ls_df_queue$df_queue_il_wide
df_alloc_i_long <- ls_df_queue$df_alloc_i_long
df_rho_gini  <- ls_df_queue$df_rho_gini
df_alloc_il_long <- ls_df_queue$df_alloc_il_long
```

Show output dataframes:

```{r}
# Display Results
print(head(df_queue_il_long, 10))
# print(str(df_queue_il_long))
print(head(df_queue_il_wide, 10))
# print(str(df_queue_il_wide))
print(head(df_alloc_i_long, 10))
# print(str(df_alloc_i_long))
print(head(df_alloc_il_long, 10))
```

Show GINI results: 

```{r}
print(head(df_rho_gini, 30))
```

### Aggregate Statistics Outcome Grapher 

Create a function to facilitate graphing over various outcomes of interest:

```{r}
ffi_plot_agg_stats <- function(df_rho_gini, svr_agg_outcome='gini_EH_star'){
  # All figures share the same x-axis, just plot out different aggregate y outcomes
  
  x.labels <- c('lambda=0.99', 'lambda=0.90', 'lambda=0', 'lambda=-10', 'lambda=-100')
  x.breaks <- c(0.01, 0.10, 1, 10, 100)
  
  pl_out <- df_rho_gini %>%
    mutate(one_minus_rho = 1 - rho_val) %>%
    ggplot(aes(x=one_minus_rho, y=!!sym(svr_agg_outcome))) +
    geom_line() + geom_point() +
    scale_x_continuous(trans='log10', labels = x.labels, breaks = x.breaks) +
    theme_bw(base_size=8)  
  
  return(pl_out)
}
```


#### Atkinson Graph

```{r}
pl_out_atkinson_EH_star <- ffi_plot_agg_stats(df_rho_gini, svr_agg_outcome='atkinson_EH_star')
pl_out_atkinson_EH_star
```

#### GINI of Allocations Graph

```{r}
pl_out_gini_D_star <- ffi_plot_agg_stats(df_rho_gini, svr_agg_outcome='gini_D_star')
pl_out_gini_D_star
```

#### GINI of Outcome Graph

```{r}
pl_out_gini_EH_star <- ffi_plot_agg_stats(df_rho_gini, svr_agg_outcome='gini_EH_star')
pl_out_gini_EH_star
```

#### Minimum outcomes

Should be increasing with x-axis.

```{r}
pl_out_min_EH_star <- ffi_plot_agg_stats(df_rho_gini, svr_agg_outcome='min_EH_star')
pl_out_min_EH_star
```

#### Average Outcomes

Should be decreasing with x-axis.

```{r}
pl_out_mean_EH_star <- ffi_plot_agg_stats(df_rho_gini, svr_agg_outcome='mean_EH_star')
pl_out_mean_EH_star
```

#### SD of Outcomes

```{r}
pl_out_sd_EH_star <- ffi_plot_agg_stats(df_rho_gini, svr_agg_outcome='sd_EH_star')
pl_out_sd_EH_star
```


## REV Compute and Graph

### Compute REV 

```{r}
tb_rho_rev <-
  PrjOptiAlloc::ffp_opt_anlyz_sodis_rev(
    ar_rho,
    fl_teacher_increase_number,
    df_input_ib = df_opt_caschool_input_ib,
    df_queue_il_long_with_V = df_queue_il_long)
```

```{r}
# Display Results
print(tb_rho_rev)
```

### REV Graph

```{r}
x.labels <- c('lambda=0.99', 'lambda=0.90', 'lambda=0', 'lambda=-10', 'lambda=-100')
x.breaks <- c(0.01, 0.10, 1, 10, 100)
tb_rho_rev %>%
  mutate(one_minus_rho = 1 - rho_val) %>%
  mutate(REV = 100*REV) %>%
  ggplot(aes(x=one_minus_rho, y=REV)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(trans='log10', labels = x.labels, breaks = x.breaks) +
  theme_bw(base_size=8) +
  ylim(0, 100)
```

