---
title: INTRO--Optimal Allocation Two Individuals Example
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{INTRO--Optimal Allocation Two Individuals Example}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  urlcolor: blue
---

# Optimal Allocation Between Two Individuals

Determine how to optimally allocate resources between two individuals. There are two structures:

1. Allocating bounded continuous resources between two indivdiuals, where the effects of resources could be heterogeneous and the costs of resources could differ. This problem is solved by the [bounded continuous optimal allocation function](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin.html)
2. Allocating discrete resources between two individuals, where the effects of each additional unit on each individual is different and the two resources have the same cost. This problem is solved by the [discrete optimal allocation function](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin_dis.html).

For both problems, we cans solve the optimal allocation problem given different levels of aggregate resources, different planner inequality aversion parameters as well as biases.

The goal below is to generate the data input structure for these two problems in a comparable way. Generate problem solutions by calling the discrete and continuous optimal functions we have created, and generate additional outputs needed to visualize and understand the model solution. The goal is to link the results here to the basics of 2 dimensional indifference and resource sets, and illustrate how the solutions here complete each other, and add to existing results.


```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tibble)
library(forcats)
library(ggplot2)

library(REconTools)
library(PrjOptiAlloc)

library(knitr)
library(kableExtra)

library(tidyr)
library(dplyr)

# spn_file <- 'C:/Users/fan/PrjOptiAlloc/vignettes/ffv_opt_n2_fn'
# knitr::purl(paste0(spn_file, '.Rmd'), output=paste0(spn_file, ".R"), documentation = 1)
```

## Input and Output

This will be written as a function so that a great variety of cases and situations can be simulated and drawn. Intuitively, on the 2D surface, our objective is to show that choices are constrained in choice windows (choice boxes), there is a choice strike zone, in the continuous case, it is a rectangle, in the discrete case, it is a number of possibly uneven points. There will be a rightward expanding line inside this box, that is the income expansion path. In the discrete case, it is not a line, but traces through a set of points. In the continuous case, it is a line with possible kink (only no kink if disects the diagnol of the choice strike zone exactly). Each point along this line is associated with a different level of value, which is utility given optimal choices. 

Note that when marginal effects are linear, strike zone budget lines are not straight but backward bending. 

The input for strike zone are:

1. $A_i$
2. $\alpha_i$, which stands in for price as well, $\frac{\alpha_i}{p_i}$
3. $\alpha_{il}$, which generates $A_{il}$ given $A_i$

The inputs for the expansion expansion path additional includes:

1. $\left\{\lambda\right\}_q$ some vector of preference
2. $\left\{W\right\}_q$ some vector of price

The outputs of the structure here includes:

1. Continuous: A vector of y-intercepts for the continuous strike zone, and one single slope
2. Discrete: all choice combination points
3. Vector of $\left\{x^{\ast},y^{\ast},v^{\ast}\right\}_{w, \lambda}$ points for optimal choices as income expands and for different planner preference levels.

## Generate Data Inputs

We want to put the continuous and discrete problems in the same scale. They already are by construction, the x and y axis are in terms of $EH_i$ outcomes. The issue is to put the allocations in the same scale. The continuous and discrete inputs should have approximately the same scale in terms of their effects on outcomes. The aggregate effects of choices on outcomes is $\mathcal{A}_i = \sum^{W^{units}}_l A_{i,l}$, the aggregate effects of the continuous linear input variable is $W^{dollar}\cdot\frac{\alpha_i}{p_i}$.

For individual $i=1,2$, one of the two individuals, we can set their total x-axis scale to be the same:

$$\sum^{W^{units}}_l A_{i,l} = W^{dollar}\cdot\frac{\alpha_i}{p_i}$$

Given the formulation above, the procedure to generate discrete and continuous consistent marginal effects parameters is:

1. Start with $W^{dollar}$, and draw $\alpha$ and $p$ from somewhat tightly bounded uniform distribution, so they are not too different. 
2. Pick any $W^{units}$, draw uniformly $A_{il}$ for both $i=1$ and $i=2$, they must sum up to the right hand side. 

Now there are $\frac{\alpha_i}{p_i}$ as well as $A_{il}$. The aggregate effects of allocating all discrete inputs for one individual, and the aggregate effects of allocating all continuous inputs for one individual on outcome for that individual are the same. This makes the discrete and continuous comparison on the same figure possible.

Write this as a function so that we can easily generate additional examples. The function should be able to output for more than two individuals, any number of individuals for testing. 

```{r opti alloc n2 generate test}
ffi_draw_n_alpha <- function(fl_w_dollar = 100,
                             it_w_units_avg = 5,
                             fl_eh_max_inc_i = 5,
                             fl_min_max_eh_inc_ratio = 1.5,
                             fl_disc_cts_match_ratio = 1.1,
                             fl_A_i_alpha_i1_ratio = 2,
                             fl_A_i_relative_ratio = 3,
                             it_N = 2,
                             it_rand_seed = 123){
  #' @param it_w_units_avg float number of goods for cts at max, approxi number for dis
  #' @param fl_eh_max_inc_i1 float maximum increase EH for one individual
  #' @param fl_min_max_eh_inc_ratio float ratio of maximum and minimum outcome increase
  #' @param it_N integer the number of individuals 
  #' @param fl_disc_cts_match_ratio float if 1, disc unit = it_w_units_avg
  #' @param fl_A_i_alpha_i1_ratio float A_i divide the alphi_il ratio, discrete set this
  
  set.seed(it_rand_seed)
  
  # Let price be the same, does not matter, only marginal effect over price ratio matters
  fl_price <- fl_w_dollar/it_w_units_avg
  fl_cts_alpha_max <- ((fl_eh_max_inc_i*fl_price)/fl_w_dollar)
  fl_cts_alpha_min <- fl_cts_alpha_max/fl_min_max_eh_inc_ratio
  
  # Generate two individuals at min and max, otherwise draw randomly between them as well 
  if (it_N == 1) {
    stop("Need to draw for at least two individuals")
  } else {
    ar_cts_alpha <- c(fl_cts_alpha_min, fl_cts_alpha_max)
    if (it_N > 2) {
      ar_cts_alpha_more <- 
        runif((it_N-2), min=fl_cts_alpha_min, max=fl_cts_alpha_max)
      ar_cts_alpha <- c(ar_cts_alpha, ar_cts_alpha_more)
    }
    ar_cts_alpha <- sort(ar_cts_alpha, decreasing=TRUE)
  }
  
  # Aggregate W for Discrete
  ar_A_i_sum <- fl_w_dollar*(ar_cts_alpha/fl_price)
  
  # Draw A_il 
  ar_it_w_units <- sample(round(
    seq(it_w_units_avg/fl_disc_cts_match_ratio, 
        it_w_units_avg*fl_disc_cts_match_ratio, length.out=it_N)), 
    it_N, replace=TRUE)
  
  # Genereate alpha_il
  ls_ar_alpha_il = apply(
    cbind(ar_it_w_units, ar_A_i_sum),
    1,
    function(row, min, max) {
      it_draw <- row[1]
      fl_sum <- row[2]
      ar_unif <- runif(it_draw, 
                       min=fl_cts_alpha_min,
                       max=fl_cts_alpha_max) 
      ar_unif <- sort(ar_unif, decreasing=TRUE)
      ar_share <- ar_unif/sum(ar_unif)
      ar_levels <- (ar_share*fl_sum*fl_disc_cts_match_ratio)
      ar_share <- ar_share[ar_levels < fl_sum]
      ar_levels <- ar_levels[ar_levels < fl_sum]
      return(list(ar_share=ar_share,
                  ar_levels=ar_levels))
    })
  
  # Print All
  message(ls_ar_alpha_il)
  message(fl_w_dollar/fl_price)
  sapply(seq(it_N), function(x) sum(ls_ar_alpha_il[[x]]$ar_levels))
  message((fl_w_dollar*ar_cts_alpha)/fl_price)
  
  # Generate A_i that are consistent with alpha_i scales
  # randomly draw two individual N2 where the differences is 2x of A_il
  # Average of alpha_i1 of all individuals
  fl_alpha_i1_avg = mean(sapply(
    seq(it_N), function(x) ls_ar_alpha_il[[x]]$ar_levels[1]))
  # Average of A_i
  # C:\Users\fan\R4Econ\math\solutions\fs_solu_x_lin.Rmd
  fl_a <- fl_A_i_relative_ratio
  fl_b <- fl_alpha_i1_avg
  fl_A_i_gap = (fl_b*(fl_a - 1))/(1 + fl_a)
  ar_A_i = runif(it_N,min=fl_b - fl_A_i_gap, max=fl_b + fl_A_i_gap)
  
  # return otuputs
  return(list(ar_A_i = ar_A_i,
              ls_ar_alpha_il=ls_ar_alpha_il,
              fl_price = fl_price,
              ar_cts_alpha = ar_cts_alpha,
              it_rand_seed=it_rand_seed))
}

# Call Function
# ls_out_n4 <- suppressMessages(ffi_draw_n_alpha(it_N = 4))
# if called function without seed first, it_rand_seed fixed
it_rand_seed <- sample(10^7, 1, replace=T)
fl_w_dollar = 100
it_rand_seed <- 4596553
it_N <- 2
ls_out_n2 <- suppressMessages(
  ffi_draw_n_alpha(fl_w_dollar=fl_w_dollar, 
                   it_N = it_N, it_rand_seed=it_rand_seed))
ls_out_n2
```

### Generate A and alpha

These parameters persist and are used throughout the rest of the file for allocation allocations. Parameters are more or less shared by discrete and continuous parameters.

```{r}
# Draw data
fl_beta_1 = 0.5
fl_beta_2 = 1-fl_beta_1
fl_w_dollar = 100
it_w_units_avg = 5
it_rand_seed <- sample(10^7, 1, replace=T)
it_rand_seed <- 8135788
it_N <- 2
ls_out_n2 <- suppressMessages(
  ffi_draw_n_alpha(fl_w_dollar=fl_w_dollar, 
                   it_w_units_avg=it_w_units_avg,
                   it_N = it_N, it_rand_seed=it_rand_seed))

# Translate to 2D budget indiff units
# fl_e = endowment points
fl_A_1 = ls_out_n2$ar_A_i[1]
fl_A_2 = ls_out_n2$ar_A_i[2]
# fl_alpha_1l = alpha_il 
# fl_eh_alloc_il = outcome at alloations
ar_alpha_1l = ls_out_n2$ls_ar_alpha_il[[1]]$ar_levels
ar_alpha_2l = ls_out_n2$ls_ar_alpha_il[[2]]$ar_levels
ar_alpha_1l_zr = c(0,ls_out_n2$ls_ar_alpha_il[[1]]$ar_levels)
ar_alpha_2l_zr = c(0,ls_out_n2$ls_ar_alpha_il[[2]]$ar_levels)
ar_A_1l <- fl_A_1 + cumsum(head(ar_alpha_1l_zr, -1))
ar_A_2l <- fl_A_2 + cumsum(head(ar_alpha_2l_zr, -1))
ar_eh_1l_zr <- fl_A_1 + cumsum(ar_alpha_1l_zr)
ar_eh_2l_zr <- fl_A_2 + cumsum(ar_alpha_2l_zr)
it_D_1l <- length(ar_alpha_1l)
it_D_2l <- length(ar_alpha_2l)
# Max Discrete Choice After which budget outside of choice strike zone fully
it_w_max <- length(ar_alpha_1l) + length(ar_alpha_2l)
ar_w_solve_at_disc <- seq(1, it_w_max)

# Continuous
fl_alpha_1 = ls_out_n2$ar_cts_alpha[1]
fl_alpha_2 = ls_out_n2$ar_cts_alpha[2]
fl_price = ls_out_n2$fl_price
# continuous choice use discrete max as bounds
fl_D_max_1 = length(ar_alpha_1l)
fl_D_max_2 = length(ar_alpha_2l)
fl_EH_max_1 = fl_A_1 + fl_D_max_1*fl_alpha_1
fl_EH_max_2 = fl_A_2 + fl_D_max_2*fl_alpha_2
# max_dollar 
fl_w_dollar_max = fl_price*it_w_max
ar_w_dollars_strike_at_disc <- seq(0, fl_w_dollar_max, length.out=(it_w_max))
ar_w_dollars_solve_at_disc <- seq(0, fl_w_dollar_max, length.out=100)

# print 
print(paste0('it_rand_seed:',it_rand_seed))
print(paste0('fl_A_1:', fl_A_1, ',fl_A_2:', fl_A_2))
print(paste0('fl_alpha_1:', fl_A_1, ',fl_alpha_2:', fl_A_2))
print(paste0('ar_alpha_1l:', ar_alpha_1l))
print(paste0('ar_alpha_2l:', ar_alpha_2l))
print(paste0('ar_A_1l:', ar_A_1l))
print(paste0('ar_A_2l:', ar_A_2l))
print(paste0('ar_eh_1l_zr:', ar_eh_1l_zr))
print(paste0('ar_eh_2l_zr:', ar_eh_2l_zr))
```

## Feasible Allocation and Utility

In this section, we generate all possible combinations of feasible allocation across these two individuals, and also indifference curve points combinations at each choice points. 

### Discrete Resource Exhausting Allocation Points

Given aggregate units of resources, from 1 to *it_w_units_avg*.

```{r}
ls_mt_strike_eh_disc <- vector(mode = "list", length = it_w_units_avg)

for (it_w in ar_w_solve_at_disc){
  
  ar_cur_w_ehlt_combo <- c()
  ar_cur_w_unit_combo <- c()
  
  for (it_1l in seq(1, length(ar_alpha_1l_zr))){
    fl_eh_1l <- ar_eh_1l_zr[it_1l]
    
    for (it_2l in seq(1, length(ar_alpha_2l_zr))){
      fl_eh_2l <- ar_eh_2l_zr[it_2l]
      
      # included alloca = 0 at L162
      it_1l_2l_sum <- (it_1l-1) + (it_2l-1)
      if (it_1l_2l_sum == it_w){
        # at this point, EH_1 and EH_2
        
        # Aggregate Resource is Fully Distributed
        ar_cur_w_ehlt_combo <- c(ar_cur_w_ehlt_combo, fl_eh_1l, fl_eh_2l)
        ar_cur_w_unit_combo <- c(ar_cur_w_unit_combo, it_1l-1, it_2l-1)
        
      }
    }
  }
  
  dim(ar_cur_w_ehlt_combo) <- c(2, length(ar_cur_w_ehlt_combo)/2)
  dim(ar_cur_w_unit_combo) <- c(2, length(ar_cur_w_unit_combo)/2)
  mt_strike_eh_disc <- cbind(it_w*fl_price, it_w, 
                             t(ar_cur_w_ehlt_combo), 
                             t(ar_cur_w_unit_combo))
  colnames(mt_strike_eh_disc) <- 
    c('w_dollars', 'w_units', 'EH_1l', 'EH_2l', 'D_1l', 'D_2l')
  ls_mt_strike_eh_disc[[it_w]] <- mt_strike_eh_disc
  
}

tb_strike_eh_disc_all <- 
  as_tibble(do.call(rbind, ls_mt_strike_eh_disc)) %>% 
  mutate(vartype = 'discrete') %>% 
  mutate(outtype = 'strike')

print(tb_strike_eh_disc_all)
```


### Continuous Choice Strike Zone

The continuous choice strike zone for the CRS problem with bounded choices is a box with straight downward sloping lines through representing budget lines. All lines are parallel to each other. We are interseted in the intercepts and slopes. Slopes are all the same, so just the intercepts. There is a smooth set of intercepts, but for graphing etc, we only need a finite set of lines representing the discrete problem's resource points. And we need the bounding corner values. 

```{r}
# assume 1 is the x-axis and 2 is the y-axis
ffi_cts_strike_crs_slope_intercept <- 
  function(fl_w, fl_alpha_price_rat_1, fl_alpha_price_rat_2, fl_A_1, fl_A_2) {
    
    fl_slope = 
      (-1)*(fl_alpha_price_rat_2/fl_alpha_price_rat_1)
    
    fl_intercept = 
      fl_w*fl_alpha_price_rat_2 + 
      (fl_alpha_price_rat_2/fl_alpha_price_rat_1)*fl_A_1 + 
      fl_A_2
    
    return(list(fl_slope=fl_slope, fl_intercept=fl_intercept))
  }
```

Call the continuous slope generator.

```{r}
# Define Marginal Effects and Price Ratios
fl_alpha_price_rat_1 <- fl_alpha_1/fl_price
fl_alpha_price_rat_2 <- fl_alpha_2/fl_price

# Continuous slope generator
ls_ar_draws_shares = sapply(ar_w_dollars_strike_at_disc,
                            ffi_cts_strike_crs_slope_intercept,
                            fl_alpha_price_rat_1=fl_alpha_price_rat_1, 
                            fl_alpha_price_rat_2=fl_alpha_price_rat_2,
                            fl_A_1=fl_A_1,
                            fl_A_2=fl_A_2)
mt_ar_draws_shares <- 
  cbind(ar_w_dollars_strike_at_disc, as_tibble(t(ls_ar_draws_shares)) %>% unnest()) %>% 
  rowid_to_column(var = "row_ctr")

# Print results
print(mt_ar_draws_shares)
```

A grid of x-points and corresponding y-points at each of the slopes

```{r}
ls_mt_strike_eh_cts <- 
  apply(mt_ar_draws_shares, 1,
        function(row) {
          it_row_ctr <- row[1]
          fl_w_dollars <- row[2]
          fl_slope <- row[3]
          fl_intercept <- row[4]
          ar_EH1 <- seq(fl_A_1, fl_EH_max_1, length.out=100)
          ar_EH2 <- fl_intercept + fl_slope*ar_EH1

          ar_EH1 <- ar_EH1[ar_EH2<=fl_EH_max_2]
          ar_EH2 <- ar_EH2[ar_EH2<=fl_EH_max_2]

          ar_EH1 <- ar_EH1[ar_EH2>=fl_A_2]
          ar_EH2 <- ar_EH2[ar_EH2>=fl_A_2]
          
          ar_D_1l <- (ar_EH1 - fl_A_1)/fl_alpha_1
          ar_D_2l <- (ar_EH2 - fl_A_2)/fl_alpha_2
        
          fl_w_units <- (fl_w_dollars/fl_price)
          mt_strike_eh_cts <- 
            cbind(fl_w_dollars, fl_w_units, 
                  ar_EH1, ar_EH2, ar_D_1l, ar_D_2l)
          rownames(mt_strike_eh_cts) <- NULL
          
          colnames(mt_strike_eh_cts) <- 
            c('w_dollars', 'w_units', 
              'EH_1l', 'EH_2l', 'D_1l', 'D_2l')
          
          
          return(mt_strike_eh_cts)
        })

tb_strike_eh_cts_all <- 
  as_tibble(do.call(rbind, ls_mt_strike_eh_cts)) %>% 
  mutate(vartype = 'continuous') %>% 
  mutate(outtype = 'strike')

print(head(tb_strike_eh_cts_all, 10))
```

## Optimal Choices

```{r}
fl_dis_w <- (it_w_units_avg + 1)
fl_cts_w <- fl_dis_w # common price
ar_rho <- c(0.5, -5)
ar_rho <- c(0.99, -0.01, -100)
ar_rho <- c(0.99, -0.01, -6)
# ar_rho <- 1 - (10^(c(seq(-2,2, length.out=20))))
```

### Optimal Discrete Choices

Given $A_i$ and $\alpha_{il}$, solve for optimal discrete allocation

Build up discrete input frame
```{r}
# ID and max Discrete Allocation
mt_i_D <- cbind(c(1,2), c(it_D_1l, it_D_2l))
colnames(mt_i_D) <- c('id_i', 'D_max_i')      
tb_i_D <- as_tibble(mt_i_D)
# A_i and alpha_il as matrix
mt_A_alpha <- rbind(cbind(1, seq(1,it_D_1l), ar_alpha_1l, ar_A_1l, fl_beta_1),
                    cbind(2, seq(1,it_D_2l), ar_alpha_2l, ar_A_2l, fl_beta_2))
colnames(mt_A_alpha) <- c('id_i', 'D_il', 'alpha_il', 'A_il', 'beta_i')
# Combine to generate input_il matrix
df_input_il <- tb_i_D %>% 
  uncount(D_max_i) %>% 
  rowid_to_column(var = "id_il") %>% 
  left_join(tb_i_D, by='id_i') %>%
  group_by(id_i) %>%
  mutate(D_il = row_number()) %>%
  left_join(as_tibble(mt_A_alpha), by=(c('id_i'='id_i', 'D_il'='D_il')))
```

Solve for optimal choices.

```{r}
ls_dis_solu <- suppressWarnings(suppressMessages(
  ffp_opt_anlyz_rhgin_dis(ar_rho,
                          fl_dis_w,
                          df_input_il,
                          bl_df_alloc_il = TRUE,
                          bl_return_V = TRUE,
                          bl_return_allQ_V = TRUE,
                          bl_return_inner_V = TRUE)))
df_queue_il_long_n2 <-ls_dis_solu$df_queue_il_long
df_queue_il_wide_n2 <- ls_dis_solu$df_queue_il_wide
df_alloc_i_long_n2 <- ls_dis_solu$df_alloc_i_long
df_rho_gini_n2 <- ls_dis_solu$df_rho_gini
df_alloc_il_long_n2 <- ls_dis_solu$df_alloc_il_long
```

Solve and select only information relevant for showing how choices change along preference specific income expansion paths. This is very wasteful, but it's OK. I only output optimal choices in the alloc_i matrix, which is resource specific. So to get the proper income expansion results, resolve at each element of resource from min to max

```{r}
ls_mt_opti_eh_disc <- apply(cbind(ar_w_solve_at_disc), 1,
  function(row){
    mt_opti_disc <- suppressWarnings(suppressMessages(
    ffp_opt_anlyz_rhgin_dis(
      ar_rho,
      row[1],
      df_input_il,
      bl_df_alloc_il = TRUE,
      bl_return_V = TRUE,
      bl_return_allQ_V = TRUE,
      bl_return_inner_V = TRUE)$df_alloc_i_long)) %>%
    select(rho_val, D_star_i, EH_star_i) %>%
    pivot_wider(names_from = id_i,
                values_from = c(D_star_i, EH_star_i)) %>%
    rename(EH_1l = EH_star_i_1, EH_2l = EH_star_i_2, 
           D_1l = D_star_i_1, D_2l = D_star_i_2) %>%
    mutate(w_units = row[1], w_dollars = row[1]*fl_price) %>%
    select(w_dollars, w_units, EH_1l, EH_2l, D_1l, D_2l, rho_val)
  return(mt_opti_disc)
  })
tb_opti_eh_disc_all <- 
  as_tibble(do.call(rbind, ls_mt_opti_eh_disc)) %>% 
  mutate(vartype = 'discrete') %>% 
  mutate(outtype = 'optimal')
print(head(tb_opti_eh_disc_all, 10))
```

### Optimal Continuous Choices

Now I take advantage of the bounded linear solution and solve for the continuous allocation problem given the previously generated continuous parameters that are consistent (in the same general scale) as the discrete parameters used for the discrete allocations. 

Build up continuous input frame

```{r}
# ID and max Discrete Allocation
# Note the problem here does not explicity 
mt_cts <- cbind(c(1,2), 
                c(fl_A_1, fl_A_2),
                c(fl_alpha_1, fl_alpha_2),
                c(fl_beta_1, fl_beta_2))
colnames(mt_cts) <- c('id_i', 'A_i', 'alpha_i', 'beta_i')      
tb_cts <- as_tibble(mt_cts)
```

```{r}
df <- tb_cts
svr_id_i <- 'id_i'
svr_A_i <- 'A_i'
svr_alpha_i <- 'alpha_i'
svr_beta_i <- 'beta_i'
fl_N_agg <- fl_cts_w/fl_price
ls_lin_solu_all_rhos <- ffp_opt_anlyz_rhgin(df, svr_id_i, svr_A_i, svr_alpha_i, svr_beta_i, fl_N_agg, ar_rho)
df_opti_alloc_all_rho_n2 <- ls_lin_solu_all_rhos$df_opti_alloc_all_rho
mt_opti_alloc_all_rho_n2 <- ls_lin_solu_all_rhos$mt_opti_alloc_all_rho
mt_expc_outcm_all_rho_n2 <- ls_lin_solu_all_rhos$mt_expc_outcm_all_rho
mt_gini_n2 <- ls_lin_solu_all_rhos$mt_gini
```

Combine results so that the continuous optimal choices' outputs are in the same dataframe structure as discrete and continuous strike zone choice sets and the discrete optimal solution points

```{r}
ls_mt_opti_eh_cts <- apply(
  cbind(ar_w_dollars_solve_at_disc), 1,
  function(row){
    
    fl_w_dollars <- row[1]
    fl_w_units <- fl_w_dollars/fl_price
    
    ls_lin_solu_all_rhos <- suppressWarnings(suppressMessages(
      ffp_opt_anlyz_rhgin(
      df, svr_id_i, svr_A_i, svr_alpha_i, svr_beta_i, fl_w_units, ar_rho)))
    mt_opti_alloc_all_rho_n2 <- ls_lin_solu_all_rhos$mt_opti_alloc_all_rho
    mt_expc_outcm_all_rho_n2 <- ls_lin_solu_all_rhos$mt_expc_outcm_all_rho
    
    mt_opti_cts <- cbind(fl_w_dollars, fl_w_units, 
                         t(mt_expc_outcm_all_rho_n2), t(mt_opti_alloc_all_rho_n2),
                         ar_rho)
    colnames(mt_opti_cts) <- c('w_dollars', 'w_units', 
                               'EH_1l', 'EH_2l', 'D_1l', 'D_2l', 'rho_val')

    return(as_tibble(mt_opti_cts))
  })

tb_opti_eh_cts_all <- 
  as_tibble(do.call(rbind, ls_mt_opti_eh_cts)) %>% 
  mutate(vartype = 'continuous') %>% 
  mutate(outtype = 'optimal')
print(head(tb_opti_eh_cts_all, 10))
```

#### Bound Continuous Choices

There are hard bounds on the level of allocations allowed at most for the two individuals. If allocation max exceeded for person $i$ but not person $j$, then distribute allocation from $i$ to $j$.

1. generate new *D* variable for amount that exceeds bounds. 
2. reset the excess allocation columns so that excess with existing does not exceed individual bound, cross $i$ and $j$
3. apply function to add to existing outcome additional allocation multiplied by marginal effects. 

```{r}
# 1. generate new *D* variable for amount that exceeds bounds. 
tb_opti_eh_cts_all <- tb_opti_eh_cts_all %>%
  mutate(D_1l_excess = 
           case_when(D_1l > fl_D_max_1 
                     ~ (D_1l - fl_D_max_1),
                     TRUE 
                     ~ 0),
         D_2l_excess = 
           case_when(D_2l > fl_D_max_2 
                     ~ (D_2l - fl_D_max_2),
                     TRUE 
                     ~ 0)) %>%
  select(w_dollars, EH_1l, EH_2l, D_1l, D_2l, 
         D_1l_excess, D_2l_excess, everything())

# 2. if both exceed, then reset outcomes for both
tb_opti_eh_cts_all <- tb_opti_eh_cts_all %>%
  mutate(EH_1l = 
           case_when(D_1l_excess > 0 & D_2l_excess > 0 
                     ~ EH_1l - D_1l_excess*(fl_alpha_1),
                     TRUE ~ EH_1l),
         EH_2l = 
           case_when(D_1l_excess > 0 & D_2l_excess > 0 
                     ~ EH_2l - D_2l_excess*(fl_alpha_2),
                     TRUE ~ EH_2l))

# 3. 1 exceeds but 2 does not
tb_opti_eh_cts_all <- tb_opti_eh_cts_all %>%
  mutate(EH_1l = 
           case_when(D_1l_excess > 0 & D_2l_excess == 0 
                     ~ EH_1l - D_1l_excess*(fl_alpha_1),
                     TRUE ~ EH_1l),
         EH_2l = 
           case_when(D_1l_excess > 0 & D_2l_excess == 0 
                     ~ EH_2l + D_1l_excess*(fl_alpha_2),
                     TRUE ~ EH_2l))
# %>%
#   mutate(EH_2l =
#            case_when(EH_2l >= fl_EH_max_2 ~ fl_EH_max_2,
#                      TRUE ~ EH_2l))
         
# 4. 2 exceeds but 1 does not
tb_opti_eh_cts_all <- tb_opti_eh_cts_all %>%
  mutate(EH_1l = 
           case_when(D_1l_excess == 0 & D_2l_excess > 0 
                     ~ EH_1l + D_2l_excess*(fl_alpha_1),
                     TRUE ~ EH_1l),
         EH_2l = 
           case_when(D_1l_excess == 0 & D_2l_excess > 0 
                     ~ EH_2l - D_2l_excess*(fl_alpha_2),
                     TRUE ~ EH_2l)) %>% 
  # mutate(EH_1l = 
  #          case_when(EH_1l > fl_EH_max_1 ~ fl_EH_max_1,
  #                    TRUE ~ EH_1l)) %>%   
  select(-D_1l_excess, -D_2l_excess)
```

### Combine Discrete and Continuous Optimal Choices Evaluate Value

Combine the constrained choice sets and optimal choices together. Then evaluate utility at optimal choices

```{r}
# Combine optimal choices as well as choicesets
tb_all_points <- 
  rbind(tb_strike_eh_disc_all %>% mutate(rho_val = NA),
        tb_opti_eh_disc_all, 
        tb_strike_eh_cts_all %>% mutate(rho_val = NA),
        tb_opti_eh_cts_all) 

# Evaluate to obtain Value at Optimal Choices and Utility at Other
# choice points. 
ffi_util_eval <- function(EH_1l, EH_2l, rho_val) {
  fl_util = (fl_beta_1*(EH_1l^rho_val) + fl_beta_2*(EH_2l^rho_val))^(1/rho_val)
  return(fl_util)
}
tb_all_points <- tb_all_points %>% 
  mutate(val_at_choices = ffi_util_eval(EH_1l, EH_2l, rho_val))

# Summary results
summary(tb_all_points)
```

#### Generate Indifference Curves at Selected Points

```{r}
# A as x-axis, need bounds on A
fl_A_min = 0.01
fl_A_max = it_D_1l+it_D_1l/2
it_A_grid = 10000

# Define parameters
# ar_lambda <- 1 - (10^(c(seq(-2,2, length.out=3))))
ar_beta <- fl_beta_1

# Select v_points 
# Indifference at three points, min, max and middle of allocations
mt_v_star_rho_val <- tb_all_points %>% 
  filter(outtype == 'optimal' & 
           (w_units == 5 ) 
         # & (rho_val == ar_rho[2])
         ) %>%
  select(w_units, vartype, rho_val, val_at_choices)

# Combine for inputs to indiff function
tb_pref <- as_tibble(mt_v_star_rho_val) %>% 
  expand_grid(ar_beta) %>% 
  rename_all(~c('w_units','vartype','rho_val', 'vstar', 'beta')) %>% 
  rowid_to_column(var = "indiff_id") %>% 
  select(indiff_id, rho_val, beta, vstar, w_units, vartype)

# Generate indifference points with apply and anonymous function
# note must only select the numeric arrays
ls_df_indiff <- apply(
  tb_pref %>% 
    select(indiff_id, rho_val, beta, vstar), 
  1, function(x){
    indiff_id <- x[1]
    lambda <- x[2]
    beta <- x[3]
    vstar <- x[4]
    
    ar_fl_A_indiff <- seq(fl_A_min, fl_A_max, length.out=it_A_grid)
    ar_fl_B_indiff <- (((vstar^lambda) - 
                          (beta*ar_fl_A_indiff^(lambda)))/(1-beta))^(1/lambda)
    mt_A_B_indiff <- cbind(indiff_id, lambda, beta, vstar,
                           ar_fl_A_indiff, ar_fl_B_indiff)
    colnames(mt_A_B_indiff) <- c('indiff_id', 'rho_val', 'beta_1', 'val_at_choices',
                                 'EH_1l', 'EH_2l')
    tb_A_B_indiff <- as_tibble(mt_A_B_indiff) %>% 
      rowid_to_column(var = "A_grid_id") %>% 
      filter(EH_2l >= 0 & EH_2l <= max(ar_fl_A_indiff))
    return(tb_A_B_indiff)
  })
df_indiff <- do.call(rbind, ls_df_indiff) %>% drop_na()
df_indiff <- df_indiff %>% 
  left_join(tb_pref %>% select(indiff_id, w_units, vartype), by='indiff_id')
```

## Graph Choices Strike Set and Optimal Choices

To prepare for graphing, generat varous factors.

```{r}
tb_all_points <- tb_all_points %>% 
  mutate(rho_val = case_when(is.na(rho_val) ~ 9999, TRUE ~ rho_val)) %>%
  mutate(vartype = as_factor(vartype), 
         outtype = as_factor(outtype),
         rho_val = as_factor(rho_val)) %>%
  mutate(varouttype = fct_cross(vartype, outtype, sep='_')) 
```
### Graph Discrete and Continuous Choice Sets

Graph Continuous Results

```{r}
# Labeling
st_title <- paste0('Expected Outcomes (Continuous + Discrete Choices)')
st_subtitle <- paste0('Additional discrete inputs have i and l specific non-increasing effects\n',
                      'Continuous choices assumed to have i specific linear effects')
st_caption <- paste0('Expected Outcome Two People Example')
st_x_label <- 'Expected Outcome Individual 1'
st_y_label <- 'Expected Outcome Individual 2'

# Graphing
plt_strike <- 
  tb_all_points %>% filter(outtype=='strike') %>%
  ggplot(aes(x=EH_1l, y=EH_2l, colour=vartype, shape=vartype)) +
  geom_jitter(size=4, width = 0.01) +
  labs(title = st_title, subtitle = st_subtitle,
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# show
print(plt_strike)
```

### Graph Discrete Choice Set and Income Expansion Path. 

Income expansion given different planner preferences. 

```{r}
# Labeling
st_title <- paste0('Expected Outcomes (Discrete Choices + Optimal)')
st_subtitle <- paste0('Planner Preference and Varying Discrete Income Expansion Path')
st_caption <- paste0('Expected Outcome Two People Example')
st_x_label <- 'Expected Outcome Individual 1'
st_y_label <- 'Expected Outcome Individual 2'

# Graphing
plt_disc <- 
  tb_all_points %>% 
  filter(vartype=='discrete') %>%
  ggplot(aes(x=EH_1l, y=EH_2l, 
             colour=rho_val, shape=rho_val)) +
  geom_jitter(size=7, width = 0) +
  labs(title = st_title, subtitle = st_subtitle, 
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# show
print(plt_disc)
```

### Graph Continuous Choice Set and Income Expansion Path

Continuous choice set and income expansion path. Unlike in the discrete case, each additional increment has the same effect for the same individual. This is assuming linear effects. Alternatively, could assume the effects are decreasing. 


```{r}
# Labeling
st_title <- paste0('Expected Outcomes (Continuous Choices + Optimal)')
st_subtitle <- paste0('Planner Preference and Varying Continuous Income Expansion Path\n',
                      'Assume marginal effects are heterogeneous for i but constant across l')
st_caption <- paste0('Expected Outcome Two People Example')
st_x_label <- 'Expected Outcome Individual 1'
st_y_label <- 'Expected Outcome Individual 2'

# Graphing
plt_cts <-
  tb_all_points %>% 
  filter(vartype=='continuous') %>%
  ggplot(aes(x=EH_1l, y=EH_2l, 
             colour=rho_val, shape=rho_val)) +
  geom_jitter(size=5, width = 0) +
  labs(title = st_title, subtitle = st_subtitle, 
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# show
print(plt_cts)
```


### Jointly Graph Optimal Discrete and Continuous Choices


```{r}
# Labeling
st_title <- paste0('Optimal Choices (Continuous + Discrete)')
st_subtitle <- paste0('Planner Preference and Varying Continuous Income Expansion Path\n',
                      'Assume marginal effects are heterogeneous for i but constant across l\n',
                      'Discrete and Continuous Problems are kind of on the same scale for comparison')
st_caption <- paste0('Expected Outcome Two People Example')
st_x_label <- 'Expected Outcome Individual 1'
st_y_label <- 'Expected Outcome Individual 2'

# Graphing
plt_opti <-
  tb_all_points %>% 
  filter(outtype=='optimal') %>%
  ggplot(aes(x=EH_1l, y=EH_2l, 
             colour=rho_val, shape=vartype)) +
  geom_jitter(size=2, width = 0) +
  labs(title = st_title, subtitle = st_subtitle, 
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# show
print(plt_opti)
```

### Utility At Optimal Choices Plots

Given the previous optimal choice points that we selected: min, max and middle of feasible choice, now draw them out.

```{r}
# Labeling
st_title <- paste0('Indifference Curves Aktinson Atkinson Utility (CES)')
st_subtitle <- paste0('Each Panel Different Beta Weights lambda is inequaliyt aversion')
st_caption <- paste0('Indifference Curve 2 Individuals.',
                     '')
st_x_label <- 'Individual 1 EH'
st_y_label <- 'Individual 2 EH'

# indiff_id', 'rho_val', 'beta_1', 'val_at_choices',
#                                'EH_1l', 'EH_2l

# Graphing
plt_indiff <- 
  df_indiff %>% 
  mutate(rho_val = as_factor(rho_val),
         beta_1 = as_factor(beta_1),
         val_at_choices = as_factor(val_at_choices)) %>%
  ggplot(aes(x=EH_1l, y=EH_2l, colour=rho_val)) +
  facet_wrap(~ vartype) + 
  geom_point(size=1) +
  labs(title = st_title, subtitle = st_subtitle,
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# Graphing
plt_indiff_cts <- 
  df_indiff %>% filter(vartype == 'continuous') %>% 
  mutate(rho_val = as_factor(rho_val),
         beta_1 = as_factor(beta_1),
         val_at_choices = as_factor(val_at_choices)) %>%
  ggplot(aes(x=EH_1l, y=EH_2l, colour=rho_val)) +
  geom_point(size=1) +
  labs(title = st_title, subtitle = st_subtitle,
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# Graphing
plt_indiff_disc <- 
  df_indiff %>% filter(vartype == 'discrete') %>% 
  mutate(rho_val = as_factor(rho_val),
         beta_1 = as_factor(beta_1),
         val_at_choices = as_factor(val_at_choices)) %>%
  ggplot(aes(x=EH_1l, y=EH_2l, colour=rho_val)) +
  geom_point(size=1) +
  labs(title = st_title, subtitle = st_subtitle,
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# show
print(plt_indiff)
```

#### Discrete Utility and Choices Jointly 

```{r, widh}
plt_indiff_disc + 
  geom_jitter(data = tb_all_points %>%filter(vartype=='discrete'),
              aes(x=EH_1l, y=EH_2l, colour=rho_val, shape=rho_val),
              size=7, width = 0)
```

#### Continuous Utility and Choices Jointly 

```{r, widh}
plt_indiff_cts + 
  geom_jitter(data = tb_all_points %>%filter(vartype=='continuous'),
              aes(x=EH_1l, y=EH_2l, colour=rho_val, shape=rho_val),
              size=7, width = 0)
  
```
