---
title: DISCRETE--Discrete Provision Optimal Allocation Solution (Line by Line)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DISCRETE--Discrete Provision Optimal Allocation Solution (Line by Line)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Back to **[Fan](https://fanwangecon.github.io/)**'s Optimal Allocation Homepage **[Table of Content](https://fanwangecon.github.io/PrjOptiAllo==c/)**

# Objective

The california school dataset, the effect of teacher ratio on student scores.

# Load Packages and Data

## Load Dependencies

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)

library(PrjOptiAlloc)

library(knitr)
library(kableExtra)
```

## Policy Parameters


```{r control policy}
# 30 percent teacher at most per school, discretize floor as needed
fl_max_add_teacher_percent = 0.30
# 15 percent total additional of all teachers
fl_teacher_increase_percent = 0.15
```

## Load Data

Generate four categories by initial height and mother's education levels combinations. Note the dataset does not provide all the details we want, so we will make some assumptions. Importantly, we only have averages within district. 

1. *teachers*: is full-time-equivalent average number of teacher per district. 
2. *enrl_tot*: is the enrollment average per district


```{r Load Packages and Process Data}
# Load Data
data(df_opt_caschool)
dfca <- df_opt_caschool

# Summarize
str(dfca)
summary(dfca)
dfca %>% group_by(county) %>%
  summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)

# Modifying and labeling etc.
# School characteristics are averaged across the district, don't know school count per district, implicitly assume the number of schools per district is the same.
# Some districts have larger and others smaller sized schools, ignoring within district variations.
dfca <- dfca %>%
          mutate(id = X) %>% select(-X) %>%
          select(id, everything()) %>%
          mutate(teachers = round(teachers),
                 enrltot = round(enrltot))

# Student and Teacher Ratio, str and stravg are almost identical
# use stravg because need to use enrltot and teachers variables later
dfca <- dfca %>% mutate(stravg = enrltot/teachers)
# view(dfca %>% select(stravg, str))

# Generate Discrete Version of continuous variables
# df_opt_birthwt <- df_opt_birthwt %>%
#     mutate(momwgtLowHigh = cut(lwt,
#                                breaks=c(-Inf, 129, Inf),
#                                labels=c("LW","HW"))) %>%
#     mutate(mombirthage = cut(age,
#                                breaks=c(-Inf, 24, Inf),
#                                labels=c("young","older"))) %>%
#     mutate(ftvm3 = cut(ftv,
#                                breaks=c(-Inf, 1, 2, Inf),
#                                labels=c("novisit","1visit","morevisit")))
```

# Regression with Data

## Tabulate

```{r tabulate groups}
# Tabulate groups, how many in each group, enough for group heterogeneity in effects?
dfca %>%
  group_by(county, grspan) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = grspan, values_from = freq)
```

## Regression Testing

The allocation policy is in terms of teachers. How many teachers to add. Given the current number of teachers in the school, suppose there is some capacity limit at each school, so the number of additional teachers at each school can not exceed 35 percent of the teachers that are already there.

Our estimation is based on regressing thest score on the student teacher ratio, from which we obtain a single estimate $\alpha$.

$$
Y_i = A_i + \theta \cdot \frac{S_i}{T_i} + \epsilon
$$

What is the marginal effect of adding one more unit?

What is the effect of adding an additional teacher?

We need to translate the estimate $\theta$ here into our equation's allocation scale

$$
Y_i \left(T_i\right) = A_i + \theta \cdot \frac{S_i}{T_i} + \epsilon\\
Y_i \left(T_i + 1\right) = A_i + \theta \cdot \frac{S_i}{T_i + 1} + \epsilon\\
EY_i \left(T_i + 1\right) - EY_i \left(T_i\right) = \theta \cdot S_i \cdot \left( \frac{1}{T_i + 1} - \frac{1}{T_i} \right)
$$

```{r Regression Testing}
attach(dfca)

# Math, English, and Overall and str = student teacher ratio
summary(lm(mathscr ~ str))
summary(lm(readscr ~ str))
summary(lm(testscr ~ str))
summary(lm(testscr ~ stravg))

# Regress test score on str with student and teacher counts
summary(lm(testscr ~ enrltot + teachers + str))
summary(lm(testscr ~ enrltot + teachers + stravg))

# Regress test score on str with covariates and county fe
summary(lm(testscr ~ factor(county) + calwpct + mealpct + computer + str - 1))
summary(lm(testscr ~ factor(county) + calwpct + mealpct + computer + stravg - 1))
```

## Discrete Regression

Need to convert regression results to increment effect.

```{r}
# Store Regression Results
mt_model <- model.matrix( ~ factor(county) + calwpct + mealpct + computer + stravg)
rs_scr_on_str = lm(testscr ~ mt_model - 1)
print(summary(rs_scr_on_str))
rs_scr_on_str_tidy = tidy(rs_scr_on_str)
rs_scr_on_str_tidy
```

## Construct Input Arrays $A_i$

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Post Regression Input Processing}
# Estimates Table
head(rs_scr_on_str_tidy, 6)
# Covariates
head(mt_model, 5)

# Covariates coefficients from regression (including constant)
ar_fl_cova_esti <- as.matrix(rs_scr_on_str_tidy %>% filter(!str_detect(term, 'stravg')) %>% select(estimate))
ar_fl_main_esti <- as.matrix(rs_scr_on_str_tidy %>% filter(str_detect(term, 'stravg')) %>% select(estimate))
head(ar_fl_cova_esti, 5)
head(ar_fl_main_esti, 5)

# Select Matrix subcomponents
mt_cova <- as.matrix(as_tibble(mt_model) %>% select(-contains("stravg")))
# mt_intr <- model.matrix(~ factor(race) - 1)
mt_intr <- t(t(rep(1, dim(mt_model)[1])))

# Generate A_i, use mt_cova_wth_const
ar_A_m <- mt_cova %*% ar_fl_cova_esti
head(ar_A_m, 5)

# Generate temporary theta_i (alpha elswhere, but theta here, because how student count will enter
ar_theta_m <- mt_intr %*% ar_fl_main_esti
head(ar_theta_m, 5)

# Beta
ar_beta_m <- rep(1/length(ar_A_m), times=length(ar_A_m))

# Dataframe of A and temporary theta
mt_A_theta_beta <- cbind(ar_A_m, ar_theta_m, ar_beta_m)

colnames(mt_A_theta_beta) <- c('A_i', 'theta_i', 'beta_i')
```

# Construct Increment Specific Effect of Each Additional Teacher

## Individual alpha_i

So this task is less trivial to do here. The procedure will be the following,

### Generate Expanded Long Frame

**Step 0**

*mt_ori*: this is the original dataframe with all variables, each row is a different observation

**Step 1**

*mt_a*: Six column matrix: col 1, district index, col 2, max number of additional teachers each district could acquire (which is determined by the number of existing teachers). There could be a maximum and minimum overall, 0 and say 100 additional teacers per district at most regardless. And each district gets up to potentially 25 percent or 30 percent additional teachers. Col 3, the number of current level of teachers in the district. Col 4, the number of current number of students. Col 5, the A_i without the addition of teacher student ratio. Col 6, not alpha_i, but just the coefficient theta that might be already individual specific due to earlier regression allowing for interaction, the alpha_i which is the $\theta_i$ possibly heterogeneous effects of str on outcomes of interest.

**Step 2**

*mt_b*: expand *mt_a*. Each row is a feasible additional teacher for a district, each there are now five variables, ID, previous level of teachers, and new level of teachers, and the existing level of students, and original A_i.

```{r generate longer dataframe}
# dfca_exmain: dfca expand main frame, five variables
dfca_exmain <- dfca %>%
                mutate(D_i = floor(fl_max_add_teacher_percent*teachers)) %>%
                select(id, D_i, teachers, enrltot)

dfca_exmain <- cbind(dfca_exmain, mt_A_theta_beta)

# following: https://fanwangecon.github.io/R4Econ/summarize/count/fs_count_basics.html
# teachers is a count coloumn to expand by 
# i_d_l is each incremental teacher
# teachers_addl is existing teacher level added up to current l
dfca_exmain <- dfca_exmain %>% uncount(D_i) %>% 
                  group_by(id) %>% 
                  mutate(i_d_l = row_number()) %>% 
                  mutate(teachers_addl = teachers + i_d_l - 1)
# display
kable(dfca_exmain[0:25,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

**Step 3**

*mt_b*: add a alpha_i column, generate the marginal productivity value, and also generate actual updated A_i. Apply the formula stated earlier. 

**Step 4**

*mt_c*: merge together the matrix now has alpha_i with also the already existing A_i.

```{r marginal effect of an additional teacher here, fig.height = 4, fig.width = 6, fig.align = "center"}
# Generate alpha_i for each additional teacher
dfca_exmain <- dfca_exmain %>% mutate(alpha_il = theta_i*enrltot*((1/teachers_addl) - (1/(teachers_addl-1)))) %>%
                                   mutate(A_il = A_i + theta_i*(enrltot/(teachers_addl-1))) %>%
                                   rowid_to_column(var = "IDil")

# Graph of the Distribution of Alpha_i and A_i
scatter <- dfca_exmain %>% ggplot(aes(x=A_il)) +
      geom_point(aes(y=log(alpha_il)), size=4, shape=4, color="red") +
      geom_abline(intercept = 0, slope = 1) + # 45 degree line
      labs(title = paste0('Discrete Allocation\nCalifornia Test Score and Student Teacher Ratio\nThe Relationship between A_il and alpha_il'),
           x = 'A_il = EXAM(teacher current + n)',
           y = 'log(alpha_il = E(teacher + n + 1) - E(teacher + n))',
           caption = paste0('Stock and Watson California Test and STR')) +
      theme_bw()
      
print(scatter)


# display
kable(dfca_exmain[0:75,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


**Step 5**

Sort, and select top given total resource avaiable, The count the total allocation for each. This means to invoke the binary allocation function. 

```{r invoke the optimal targeting function}
# Optimal Targeting
ar_rho = c(-100, -0.001,  0.95)
ar_rho <- unique(ar_rho)
ls_bin_solu_all_rhos <-
  ffp_opt_anlyz_rhgin_bin(dfca_exmain, svr_id_i = 'IDil',
                          svr_A_i = 'A_il', svr_alpha_i = 'alpha_il', svr_beta_i = 'beta_i',
                          ar_rho = ar_rho,
                          svr_inpalc = 'opti_alloc_queue',
                          svr_expout = 'opti_exp_outcome')

# Results from optimal targeting
df_all_rho <- ls_bin_solu_all_rhos$df_all_rho
df_all_rho_long <- ls_bin_solu_all_rhos$df_all_rho_long

# display
kable(df_all_rho[0:500,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


Back to dataframe mt_ori, incorporate into that, but do and shift for each rho.

```{r}
# Child Weight
```
