---
title: Function Test Discrete Allocation with Linear Regression
description: |
  Test binary optimal allocation solution function from package.
  Use the California student test score dataset, same as line by line.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function Test Discrete Allocation with Linear Regression}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---


Test discrete optimal allocation solution line by line without function. Use the California student test score dataset. Regress student English and Math test scores on Student-Teacher-Ratio.


This function produces the same results as [DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Line by Line)](https://fanwangecon.github.io/PrjOptiAlloc/articles/ffv_opt_sodis_rkone_casch_allrw.html). The differences are:

1. the program here uses the dataframes created by the line by line file as Inputs
2. the program here uses functions that are identical to the line by line function, but are saved in the R folder.

This file, by invoking functions, is much shorter than the line by line program.


```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)

library(PrjOptiAlloc)

library(knitr)
library(kableExtra)
```

## Hand Input Data Example

Set Preference Array:

```{r}
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=4))))
ar_rho <- unique(ar_rho)
it_w_agg <- 5
```

```{r}
id_i    <- c(1,1,1,2,2,2,3,3,3)
id_il   <- c(1,2,3,4,5,6,7,8,9)
D_max_i <- c(3,3,3,3,3,3,3,3,3)
D_il    <- c(1,2,3,1,2,3,1,2,3)

A_il_i1 <- c(4,3,2,1)
A_il_i2 <- c(2.7,2.2,1.5,1.1)
A_il_i3 <- c(1.9,1.8,1.1,0.1)
A_il <- c(A_il_i1[2:4], A_il_i2[2:4], A_il_i3[2:4])
alpha_il <- c(diff(-A_il_i1), diff(-A_il_i2), diff(-A_il_i3))
beta_i <- c(1/3,1/3,1/3,1/3,1/3,1/3,1/3,1/3,1/3)

df_handinput_input_il <- as_tibble(cbind(id_i, id_il, D_max_i, D_il, A_il, alpha_il, beta_i))

print(df_handinput_input_il)
```


```{r}
ls_df_queue_hand <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho, it_w_agg,
                                            bl_df_alloc_il = TRUE,
                                            bl_return_V = TRUE,
                                            bl_return_allQ_V = TRUE,
                                            bl_return_inner_V = TRUE,
                                            df_input_il = df_handinput_input_il)
df_queue_il_long <- ls_df_queue_hand$df_queue_il_long
df_queue_il_wide <- ls_df_queue_hand$df_queue_il_wide
df_alloc_i_long  <- ls_df_queue_hand$df_alloc_i_long
df_rho_gini  <- ls_df_queue_hand$df_rho_gini
df_alloc_il_long <- ls_df_queue_hand$df_alloc_il_long
```

```{r}
print(df_queue_il_long)
print(df_queue_il_wide)
print(df_alloc_i_long)
print(df_rho_gini)
print(df_alloc_il_long)
```

## Load Pre-Generated Data

These are data thata were generated in the [DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Line by Line)](https://fanwangecon.github.io/PrjOptiAlloc/articles/ffv_opt_sodis_rkone_casch_allrw.html) function.

```{r load data}
# Load Data
data(df_opt_caschool_prep_i)
data(df_opt_caschool_input_il)
data(df_opt_caschool_input_ib)

# Show Variables
str(df_opt_caschool_prep_i)
str(df_opt_caschool_input_il)
str(df_opt_caschool_input_ib)

# Modifying Beta Value
# When individuals have the same weights, beta values do not matter
df_opt_caschool_input_il <- df_opt_caschool_input_il %>% mutate(beta_i = 1)
df_opt_caschool_input_ib <- df_opt_caschool_input_ib %>% mutate(beta_i = 1)
```

## Solve All Districts Discrete Optimal Allocation Across Planner Preferences

### Specify Policy Parameters

Note that *input_il* function already is based on *fl_fi_max*.

```{r control policy}
# 100 percent teacher at most per school, discretize floor as needed
# This is not a parameter that matters here, already a part of the input_il function
fl_fi_max = 1.00
# 20 percent total additional of all teachers
fl_fa_max = 0.20
# What is the number of teachers we can increase by
fl_teacher_increase_number <- sum(df_opt_caschool_prep_i$teachers)*fl_fa_max
fl_teacher_increase_number <- floor(fl_teacher_increase_number)

# Rho values to consider
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=4))))
ar_rho <- unique(ar_rho)
```

## Optimal Queue and Allocations

Use the discrete allocation function across preference [ffp_opt_anlyz_rhgin_dis.html](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin_dis.html) function. This computes optimal allocation for multiple planner inequality aversion $\lambda$ parameters at the same time. Note that in the function $\lambda=\rho$.

```{r}
# Optimal Allocation
ls_df_queue <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho,
                                       fl_teacher_increase_number,
                                       bl_df_alloc_il = TRUE,
                                       df_input_il = df_opt_caschool_input_il,
                                       svr_rho = 'rho',
                                       svr_id_i = 'id_i', svr_id_il = 'id_il',
                                       svr_D_max_i = 'D_max_i', svr_D_il = 'D_il',
                                       svr_D_star_i = 'D_star_i', svr_F_star_i = 'F_star_i',
                                       svr_inpalc = 'Q_il', svr_D_Wbin_il = 'D_Wbin_il',
                                       svr_A_il = 'A_il', svr_alpha_il = 'alpha_il', svr_beta_i = 'beta_i',
                                       svr_expout = 'opti_exp_outcome',
                                       st_idcol_prefix = 'sid_')
df_queue_il_long <- ls_df_queue$df_queue_il_long
df_queue_il_wide <- ls_df_queue$df_queue_il_wide
df_alloc_i_long <- ls_df_queue$df_alloc_i_long
df_rho_gini  <- ls_df_queue$df_rho_gini
df_alloc_il_long <- ls_df_queue$df_alloc_il_long
```

```{r}
# Display Results
print(head(df_queue_il_long, 10))
# print(str(df_queue_il_long))
print(head(df_queue_il_wide, 10))
# print(str(df_queue_il_wide))
print(head(df_alloc_i_long, 10))
# print(str(df_queue_il_wide))
print(df_rho_gini)
# print(str(df_alloc_i_long))
print(head(df_alloc_il_long, 10))
```

## REV Compute

```{r}
tb_rho_rev <-
  PrjOptiAlloc::ffp_opt_anlyz_sodis_rev(ar_rho,
                          fl_teacher_increase_number,
                          df_input_ib = df_opt_caschool_input_ib,
                          df_queue_il_long_with_V = df_queue_il_long)
```

```{r}
# Display Results
print(tb_rho_rev)
```


## Optimal Allocation Among Subset of Districts

### Select a Small Subset of Districts

To generate some graphs and more easily readable results, rather than using all school districts, select some random subset of school districts from the existing data frames.

1. Select a random subset of district IDs
2. New input df that only includes these IDs.

```{r discrete func test subset select}
# select four school districts
# cas = california schools
it_O <- 4
set.seed(1)
df_cas_prep_sub_i <- df_opt_caschool_prep_i %>% filter(teachers <= 100 & teachers >= 50)
df_cas_prep_sub_i <- df_cas_prep_sub_i[sample(dim(df_cas_prep_sub_i)[1], it_O, replace=FALSE),]
ar_cas_id_selected <- df_cas_prep_sub_i %>% pull(id_i)

# Select from il and ib only ids that are randomly selected
df_cas_input_sub_il <- df_opt_caschool_input_il %>% filter(id_i %in% ar_cas_id_selected)
df_cas_input_sub_ib <- df_opt_caschool_input_ib %>% filter(id_i %in% ar_cas_id_selected)

# Print
print(df_cas_prep_sub_i)
print(df_cas_input_sub_ib)
print(df_cas_input_sub_il)
```

Now we also have a new total number of teacher increase, no longer based on all districts. The policy here is to potentially increase each school district by *fl_fi_max*, already coded into the *input_il* file, so that can not change. The total number of available new teachers is *fl_fa_max* times the existing number of teachers in the selected districts overall

```{r discrete func subset new res}
fl_fa_max = 0.20
# What is the number of teachers we can increase by
fl_teacher_increase_sub_number <- sum(df_cas_prep_sub_i$teachers)*fl_fa_max
fl_teacher_increase_sub_number <- floor(fl_teacher_increase_sub_number)
```

### Optimal Queue, Allocations and REV

Use the discrete allocation function across preference [ffp_opt_anlyz_rhgin_dis.html](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin_dis.html) function. This computes optimal allocation for multiple planner inequality aversion $\lambda$ parameters at the same time. Note that in the function $\lambda=\rho$.

```{r}
# Rho values to consider
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=4))))
ar_rho <- unique(ar_rho)

# Optimal Allocation
ls_df_sub_queue <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho, fl_teacher_increase_sub_number,
                                           df_input_il = df_cas_input_sub_il,
                                           bl_df_alloc_il = TRUE)
df_queue_il_sub_long <- ls_df_sub_queue$df_queue_il_long
df_queue_il_sub_wide <- ls_df_sub_queue$df_queue_il_wide
df_alloc_i_long <- ls_df_sub_queue$df_alloc_i_long
df_rho_gini <- ls_df_sub_queue$df_rho_gini
df_alloc_il_long <- ls_df_sub_queue$df_alloc_il_long

# REV
ar_util_rev_loop_func <- ffp_opt_anlyz_sodis_rev(ar_rho,fl_teacher_increase_sub_number,
                                                 df_input_ib = df_cas_input_sub_ib,
                                                 df_queue_il_long_with_V = df_queue_il_sub_long)
```

Show Results

```{r}
# Display Results
print(df_queue_il_sub_long)
print(df_alloc_i_long)
print(df_rho_gini)
print(df_alloc_il_long)
print(ar_util_rev_loop_func)
```

<!-- ### Additional Statistics

1. Predicted Expected Outcomes for each i given allocations
2. Utility at each Queue rank point
3. Gini Statistics based on allocations

### Results Analysis

Core graphical/tabular illustration of results across Many Individuals

1. Mean and standard deviation of outcome, across $\lambda$, along with actual mean and standard deviation
2. Gini of Allocations and Outcomes, across $\lambda$, along with observed of intended Gini of Allocation or Outcomes
3. REV in percentages as well as in units of resources lost, across $\lambda$
4. Tabular results already exists for this. Rank Data/chart:
    - columns are $\lambda$ points
    - colors identify individuals, each color appears multiple times due to multiple allocations for each individual
    - show ranking changes as preference changes
    - highligh only subset (1) of individuals
    - Climbing chart? x axis is rank, y axis allocation for individual, but not sure why this would be useful. Perhaps useful for tabular outputs. Maybe only very small chart

How should illustrative graph be shown with only 2 individuals? -->


## Differential Mass Optimal Allocation 

Consider the population of schools is measure one, there are $N$ distinct types of schools, and the mass of each unique type is different, but they sum up to 1. If each school can have 5 additional teachers, than the mass of teachers is measure 5. 

Use the number of students per school as its mass, and introduce a new vector for the allocation function that is this mass. This is to illustrate ideas. The mass should not be the number of students, but the number of schools of the same type.

Note that under discrete allocation, there were 10840 available teachers to be allocated. But in the problem below, the idea is: 

1. suppose there is some system that currently allocates teachers by school attributes
2. we know the number of teachers allocated to each type of school 
3. we know the proportion of each type of school in the school population. 
4. given a population measure of 1 as the sum of all schools, if on average 4 teachers are allocated to each school, the total of teachers allocated is measure 4. In discrete terms, that just means, if there are in total 1000 different schools (not different unique type, just different schools), and on average each school gets 4 teachers, we are allocating 4000 teachers. 

Several things to note:

1. Get the total measure of resources used by the alternative allocation, this is not summing up just the mass of all individuals, which sums to one, but sums the measure of resources allocated given individual mass, and how much is allocated to each individual
2. Allocation queue: sort by allocation queue, and cumulative sum over individual mass, repeatedly over each incremental allocation for each individual
3. Welfare calculation: the welfare/utility is not by cumulative mass, but by probability mass for each type, this does not vary depending on how much is allocated. Allocation changes alpha and A, not the mass in utility. 

### Generate The mass Variable

The mass variable is not contained in *df_opt_caschool_prep_i*, but is contained in *df_opt_caschool_prep_i*, merge together to get that number of students as mass. This variable does not need to sum to one. Reweighting to sum to one should not change results.

```{r}
# Summarize
REconTools::ff_summ_percentiles(df_opt_caschool_input_il, bl_statsasrows=FALSE)
REconTools::ff_summ_percentiles(df_opt_caschool_input_ib, bl_statsasrows=FALSE)

# Mass of Students for IL file 
df_opt_caschool_input_mass_il <- df_opt_caschool_input_il %>% 
  left_join(df_opt_caschool_prep_i %>% 
              mutate(mass_i = enrltot/sum(enrltot)) %>%
              select(id_i, mass_i)
            , by='id_i') %>% ungroup()

# Mass of Students for IL file 
# IB file done with the Assumption that fl_fa_max% additional teachers per school
# Since mass per teacher/school same, total mass per school for ib allocation 
# is determined by 
df_opt_caschool_input_mass_ib <- df_opt_caschool_input_ib %>%
  left_join(df_opt_caschool_prep_i %>% 
              mutate(mass_i = enrltot/sum(enrltot)) %>%
              mutate(cumu_sum_group_i = mass_i*round(teachers*fl_fa_max)) %>%
              select(id_i, mass_i, cumu_sum_group_i)
            , by='id_i') %>% ungroup()

# Total mass
sum(df_opt_caschool_input_mass_il$mass_i)
sum(df_opt_caschool_input_mass_ib$cumu_sum_group_i)
# Alternative Allocation and Mass for Alternative Allocations
REconTools::ff_summ_percentiles(df_opt_caschool_input_mass_il, bl_statsasrows=FALSE)
REconTools::ff_summ_percentiles(df_opt_caschool_input_mass_ib, bl_statsasrows=FALSE)
```


Within Group Cumulative Mass Sum with Allocations:

```{r}
# Within Group Cumulative Sum of Mass
# For the Value Calculation, the A and alpha are updated to be given allocation up to that point
# that means the mass calculation is not mass at a particular unit of allocation, but mass
# for that individual group/type, up to the point. This is a within group cumulative sum, 
# Note that so for welfare and for allocation queue, in the welfare case, we need cumulative within
# individual group mass, and in the allocation queue, need point mass

df_opt_caschool_input_mass_il %>%
  arrange(id_i, D_il) %>% 
  group_by(id_i) %>% 
  arrange(D_il) %>% 
  mutate(cumu_sum_group_i = cumsum(mass_i)) %>%
  arrange(id_i, D_il) %>%
  ungroup()
```


### Allocations Given Mass of Available Resources

Analyze previous allocation not based on measure, do this only for one planner:

1. sort by queue
2. cumulative sum the mass variable: mass_cumu_queue, cumulative mass (resource allocated) by this queue point. 

Note that this is to be done inside the allocation function, after queue has been found, cumulative sum mass along the queue, and set cut-off at the total measure of resources available.

```{r}
# Cumulative Mass
df_queue_il_long_mass_cumu <- df_queue_il_long %>% filter(rho_val == ar_rho[1]) %>%
  select(-rho, -rho_val) %>%
  left_join(df_opt_caschool_prep_i %>% 
              mutate(mass = enrltot/sum(enrltot)) %>%
              select(id_i, mass), by='id_i') %>%
  arrange(Q_il) %>%
  mutate(mass_cumu_queue = cumsum(mass))
# View Around the Teacher Available Cut-off mass
df_queue_il_long_mass_cumu %>% 
  filter(Q_il >= fl_teacher_increase_number - 5 
          & Q_il <= fl_teacher_increase_number + 5)
# Summarize
REconTools::ff_summ_percentiles(df_queue_il_long_mass_cumu, bl_statsasrows=FALSE)
```

### Optimal Allocations with Mass

Same as before, but now include the mass column:

```{r}
fl_measure_teacher_increase_number <- sum(df_opt_caschool_input_mass_ib$cumu_sum_group_i)
ls_df_queue <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(
  ar_rho,
  fl_measure_teacher_increase_number,
  bl_df_alloc_il = TRUE,
  df_input_il = df_opt_caschool_input_mass_il,
  svr_rho = 'rho',
  svr_id_i = 'id_i', svr_id_il = 'id_il',
  svr_D_max_i = 'D_max_i', svr_D_il = 'D_il',
  svr_D_star_i = 'D_star_i', svr_F_star_i = 'F_star_i',
  svr_inpalc = 'Q_il', svr_D_Wbin_il = 'D_Wbin_il',
  svr_A_il = 'A_il', svr_alpha_il = 'alpha_il', 
  svr_beta_i = 'beta_i',
  svr_measure_i = 'mass_i',
  svr_expout = 'opti_exp_outcome',
  st_idcol_prefix = 'sid_')
df_queue_il_long_mass <- ls_df_queue$df_queue_il_long
df_queue_il_wide_mass <- ls_df_queue$df_queue_il_wide
df_alloc_i_long_mass <- ls_df_queue$df_alloc_i_long
df_rho_gini_mass  <- ls_df_queue$df_rho_gini
df_alloc_il_long_mass <- ls_df_queue$df_alloc_il_long
```

Analyze the Allocation Results based on measure:

```{r}
# Cumulative Mass
df_queue_il_long_mass_cumu <- df_queue_il_long_mass %>% filter(rho_val == ar_rho[1]) %>%
  select(-rho, -rho_val) %>%
  left_join(df_opt_caschool_prep_i %>% 
              mutate(mass = enrltot/sum(enrltot)) %>%
              select(id_i, mass), by='id_i') %>%
  arrange(Q_il) %>%
  mutate(mass_cumu_queue = cumsum(mass))
# View Around the Teacher Available Cut-off mass
df_queue_il_long_mass_cumu %>% 
  filter(mass_cumu_queue >= fl_measure_teacher_increase_number - 0.03
          & mass_cumu_queue <= fl_measure_teacher_increase_number + 0.03)
# Summarize
REconTools::ff_summ_percentiles(df_queue_il_long_mass_cumu, bl_statsasrows=FALSE)
```

### REV Compute With Mass

```{r}
tb_rho_rev <-
  PrjOptiAlloc::ffp_opt_anlyz_sodis_rev(ar_rho,
                          fl_measure_teacher_increase_number,
                          df_input_ib = df_opt_caschool_input_mass_ib,
                          df_queue_il_long_with_V = df_queue_il_long_mass,
                          svr_beta_i = 'beta_i', svr_measure_i = 'mass_i')
```

```{r}
# Display Results
print(tb_rho_rev)
```





