---
title: DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Function)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Function)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

# Discrete Optimal Allocation Allocate Teacher (Function)


This function produces the same results as [DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Line by Line)](https://fanwangecon.github.io/PrjOptiAlloc/articles/ffv_opt_sodis_rkone_casch_allrw.html). The differences are:

1. the program here uses the dataframes created by the line by line file as Inputs
2. the program here uses functions that are identical to the line by line function, but are saved in the R folder.

This file, by invoking functions, is much shorter than the line by line program.


```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)

library(PrjOptiAlloc)

library(knitr)
library(kableExtra)
```

## Hand Input Data Example

Set Preference Array:

```{r}
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=4))))
ar_rho <- unique(ar_rho)
it_w_agg <- 5
```

```{r}
id_i    <- c(1,1,1,2,2,2,3,3,3)
id_il   <- c(1,2,3,4,5,6,7,8,9)
D_max_i <- c(3,3,3,3,3,3,3,3,3)
D_il    <- c(1,2,3,1,2,3,1,2,3)

A_il_i1 <- c(4,3,2,1)
A_il_i2 <- c(2.7,2.2,1.5,1.1)
A_il_i3 <- c(1.9,1.8,1.1,0.1)
A_il <- c(A_il_i1[2:4], A_il_i2[2:4], A_il_i3[2:4])
alpha_il <- c(diff(-A_il_i1), diff(-A_il_i2), diff(-A_il_i3))
beta_i <- c(1/3,1/3,1/3,1/3,1/3,1/3,1/3,1/3,1/3)

df_handinput_input_il <- as_tibble(cbind(id_i, id_il, D_max_i, D_il, A_il, alpha_il, beta_i))

print(df_handinput_input_il)
```


```{r}
ls_df_queue_hand <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho, it_w_agg,
                                            bl_df_alloc_il = TRUE,
                                            bl_return_V = TRUE,
                                            bl_return_allQ_V = TRUE,
                                            bl_return_inner_V = TRUE,
                                            df_input_il = df_handinput_input_il)
df_queue_il_long <- ls_df_queue_hand$df_queue_il_long
df_queue_il_wide <- ls_df_queue_hand$df_queue_il_wide
df_alloc_i_long  <- ls_df_queue_hand$df_alloc_i_long
df_rho_gini  <- ls_df_queue_hand$df_rho_gini
df_alloc_il_long <- ls_df_queue_hand$df_alloc_il_long
```

```{r}
print(df_queue_il_long)
print(df_queue_il_wide)
print(df_alloc_i_long)
print(df_rho_gini)
print(df_alloc_il_long)
```

## Load Pre-Generated Data

These are data thata were generated in the [DISCRETE--Discrete Optimal Allocation California Teacher Student Ratio (Line by Line)](https://fanwangecon.github.io/PrjOptiAlloc/articles/ffv_opt_sodis_rkone_casch_allrw.html) function.

```{r load data}
# Load Data
data(df_opt_caschool_prep_i)
data(df_opt_caschool_input_il)
data(df_opt_caschool_input_ib)

# Show Variables
str(df_opt_caschool_prep_i)
str(df_opt_caschool_input_il)
str(df_opt_caschool_input_ib)
```

## Solve All Districts Discrete Optimal Allocation Across Planner Preferences

### Specify Policy Parameters

Note that *input_il* function already is based on *fl_fi_max*.

```{r control policy}
# 100 percent teacher at most per school, discretize floor as needed
# This is not a parameter that matters here, already a part of the input_il function
fl_fi_max = 1.00
# 20 percent total additional of all teachers
fl_fa_max = 0.20
# What is the number of teachers we can increase by
fl_teacher_increase_number <- sum(df_opt_caschool_prep_i$teachers)*fl_fa_max
fl_teacher_increase_number <- floor(fl_teacher_increase_number)

# Rho values to consider
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=4))))
ar_rho <- unique(ar_rho)
```

## Optimal Queue and Allocations

Use the discrete allocation function across preference [ffp_opt_anlyz_rhgin_dis.html](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin_dis.html) function. This computes optimal allocation for multiple planner inequality aversion $\lambda$ parameters at the same time. Note that in the function $\lambda=\rho$.

```{r}
ls_df_queue <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho,
                                       fl_teacher_increase_number,
                                       bl_df_alloc_il = TRUE,
                                       df_input_il = df_opt_caschool_input_il,
                                       svr_rho = 'rho',
                                       svr_id_i = 'id_i', svr_id_il = 'id_il',
                                       svr_D_max_i = 'D_max_i', svr_D_il = 'D_il',
                                       svr_D_star_i = 'D_star_i', svr_F_star_i = 'F_star_i',
                                       svr_inpalc = 'Q_il', svr_D_Wbin_il = 'D_Wbin_il',
                                       svr_A_il = 'A_il', svr_alpha_il = 'alpha_il', svr_beta_i = 'beta_i',
                                       svr_expout = 'opti_exp_outcome',
                                       st_idcol_prefix = 'sid_')
df_queue_il_long <- ls_df_queue$df_queue_il_long
df_queue_il_wide <- ls_df_queue$df_queue_il_wide
df_alloc_i_long <- ls_df_queue$df_alloc_i_long
df_rho_gini  <- ls_df_queue$df_rho_gini
df_alloc_il_long <- ls_df_queue$df_alloc_il_long
```

```{r}
# Display Results
print(head(df_queue_il_long, 10))
# print(str(df_queue_il_long))
print(head(df_queue_il_wide, 10))
# print(str(df_queue_il_wide))
print(head(df_alloc_i_long, 10))
# print(str(df_queue_il_wide))
print(df_rho_gini)
# print(str(df_alloc_i_long))
print(head(df_alloc_il_long, 10))
```

## REV Compute

```{r}
tb_rho_rev <-
  PrjOptiAlloc::ffp_opt_anlyz_sodis_rev(ar_rho,
                          fl_teacher_increase_number,
                          df_input_ib = df_opt_caschool_input_ib,
                          df_queue_il_long_with_V = df_queue_il_long)
```

```{r}
# Display Results
print(tb_rho_rev)
```


## Optimal Allocation Among Subset of Districts

### Select a Small Subset of Districts

To generate some graphs and more easily readable results, rather than using all school districts, select some random subset of school districts from the existing data frames.

1. Select a random subset of district IDs
2. New input df that only includes these IDs.

```{r discrete func test subset select}
# select four school districts
# cas = california schools
it_O <- 4
set.seed(1)
df_cas_prep_sub_i <- df_opt_caschool_prep_i %>% filter(teachers <= 100 & teachers >= 50)
df_cas_prep_sub_i <- df_cas_prep_sub_i[sample(dim(df_cas_prep_sub_i)[1], it_O, replace=FALSE),]
ar_cas_id_selected <- df_cas_prep_sub_i %>% pull(id_i)

# Select from il and ib only ids that are randomly selected
df_cas_input_sub_il <- df_opt_caschool_input_il %>% filter(id_i %in% ar_cas_id_selected)
df_cas_input_sub_ib <- df_opt_caschool_input_ib %>% filter(id_i %in% ar_cas_id_selected)

# Print
print(df_cas_prep_sub_i)
print(df_cas_input_sub_ib)
print(df_cas_input_sub_il)
```

Now we also have a new total number of teacher increase, no longer based on all districts. The policy here is to potentially increase each school district by *fl_fi_max*, already coded into the *input_il* file, so that can not change. The total number of available new teachers is *fl_fa_max* times the existing number of teachers in the selected districts overall

```{r discrete func subset new res}
fl_fa_max = 0.20
# What is the number of teachers we can increase by
fl_teacher_increase_sub_number <- sum(df_cas_prep_sub_i$teachers)*fl_fa_max
fl_teacher_increase_sub_number <- floor(fl_teacher_increase_sub_number)
```

### Optimal Queue, Allocations and REV

Use the discrete allocation function across preference [ffp_opt_anlyz_rhgin_dis.html](https://fanwangecon.github.io/PrjOptiAlloc/reference/ffp_opt_anlyz_rhgin_dis.html) function. This computes optimal allocation for multiple planner inequality aversion $\lambda$ parameters at the same time. Note that in the function $\lambda=\rho$.

```{r}
# Rho values to consider
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=4))))
ar_rho <- unique(ar_rho)

# Optimal Allocation
ls_df_sub_queue <- PrjOptiAlloc::ffp_opt_anlyz_rhgin_dis(ar_rho, fl_teacher_increase_sub_number,
                                           df_input_il = df_cas_input_sub_il,
                                           bl_df_alloc_il = TRUE)
df_queue_il_sub_long <- ls_df_sub_queue$df_queue_il_long
df_queue_il_sub_wide <- ls_df_sub_queue$df_queue_il_wide
df_alloc_i_long <- ls_df_sub_queue$df_alloc_i_long
df_rho_gini <- ls_df_sub_queue$df_rho_gini
df_alloc_il_long <- ls_df_sub_queue$df_alloc_il_long

# REV
ar_util_rev_loop_func <- ffp_opt_anlyz_sodis_rev(ar_rho,fl_teacher_increase_sub_number,
                                                 df_input_ib = df_cas_input_sub_ib,
                                                 df_queue_il_long_with_V = df_queue_il_sub_long)
```

Show Results

```{r}
# Display Results
print(df_queue_il_sub_long)
print(df_alloc_i_long)
print(df_rho_gini)
print(df_alloc_il_long)
print(ar_util_rev_loop_func)
```

<!-- ### Additional Statistics

1. Predicted Expected Outcomes for each i given allocations
2. Utility at each Queue rank point
3. Gini Statistics based on allocations

### Results Analysis

Core graphical/tabular illustration of results across Many Individuals

1. Mean and standard deviation of outcome, across $\lambda$, along with actual mean and standard deviation
2. Gini of Allocations and Outcomes, across $\lambda$, along with observed of intended Gini of Allocation or Outcomes
3. REV in percentages as well as in units of resources lost, across $\lambda$
4. Tabular results already exists for this. Rank Data/chart:
    - columns are $\lambda$ points
    - colors identify individuals, each color appears multiple times due to multiple allocations for each individual
    - show ranking changes as preference changes
    - highligh only subset (1) of individuals
    - Climbing chart? x axis is rank, y axis allocation for individual, but not sure why this would be useful. Perhaps useful for tabular outputs. Maybe only very small chart

How should illustrative graph be shown with only 2 individuals? -->
