---
title: Generate A and alpha Nutrition Height Test Regression Results
description: |
  Generate A and alpha based on a linear regression.
  Four categories based on initial height and mother education levels (Guatemala and Cebu dataset).
  Regress height on protein inputs allowing for heterogeneous effects for each of the four categories.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate A and alpha Nutrition Height Test Regression Results}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Generates data from the [Guatemala and Cebu dataset](https://www.sciencedirect.com/science/article/pii/S1570677X16300107) jitter/rand version, select cebu only between 0 and 24. Generates four categories based on initial height and mother education levels. Regress height on protein inputs allowing for heterogeneous effecst for each of the four categories.

## Set Up

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```

```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
# setwd('C:/Users/fan/PrjOptiAlloc')
library(dplyr)
library(tidyr)
library(broom)
library(stringr)
library(REconTools)

bl_save_rda = FALSE
```

## Get Data

Generate four categories by initial height and mother's education levels combinations.

```{r Load Packages and Process Data}
# Load Library

# Select Cebu Only
df_hw_cebu_m24 <- REconTools::df_hgt_wgt %>%
  filter(S.country == 'Cebu' & svymthRound == 24 & prot > 0 & hgt > 0) %>% drop_na()

# Generate Discrete Version of momEdu
df_hw_cebu_m24 <- df_hw_cebu_m24 %>%
    mutate(momEduRound = cut(momEdu,
                             breaks=c(-Inf, 10, Inf),
                             labels=c("MEduLow","MEduHigh"))) %>%
    mutate(hgt0med = cut(hgt0,
                             breaks=c(-Inf, 50, Inf),
                             labels=c("h0low","h0high")))

df_hw_cebu_m24$momEduRound = as.factor(df_hw_cebu_m24$momEduRound)
df_hw_cebu_m24$hgt0med = as.factor(df_hw_cebu_m24$hgt0med)

# Attach
attach(df_hw_cebu_m24)
```

## Regression with Data and Construct Input Arrays

### Linear Regression

Estimation *Production functions* or any other function. Below, Regression height outcomes on input interacted with the categories created before. This will generate category specific marginal effects.

```{r Linear Regression}
# Input Matrix
mt_lincv <- model.matrix(~ hgt0 + wgt0)
mt_linht <- model.matrix(~ sex:hgt0med - 1)

# Regress Height At Month 24 on Nutritional Inputs with controls
rs_hgt_prot_lin = lm(hgt ~ prot:mt_linht + mt_lincv - 1)
print(summary(rs_hgt_prot_lin))
rs_hgt_prot_lin_tidy = broom::tidy(rs_hgt_prot_lin)
```

### Log-Linear Regression

Now log-linear regressions, where inptut coefficient differs by input groups.

```{r Log-Linear Regression}
# Input Matrix Generation
mt_logcv <- model.matrix(~ hgt0 + wgt0)
mt_loght <- model.matrix(~ sex:hgt0med - 1)

# Log and log regression for month 24
rs_hgt_prot_log = lm(log(hgt) ~ log(prot):mt_loght + mt_logcv - 1)
print(summary(rs_hgt_prot_log))
rs_hgt_prot_log_tidy = broom::tidy(rs_hgt_prot_log)
```

### Construct Input Arrays $A_i$ and $\alpha_i$

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Post Regression Input Processing}

# Generate A_i
ar_Ai_lin <- mt_lincv %*% as.matrix(rs_hgt_prot_lin_tidy %>% filter(!str_detect(term, 'prot')) %>% select(estimate))
ar_Ai_log <- mt_logcv %*% as.matrix(rs_hgt_prot_log_tidy %>% filter(!str_detect(term, 'prot')) %>% select(estimate))

# Generate alpha_i
ar_alphai_lin <- mt_linht %*% as.matrix(rs_hgt_prot_lin_tidy %>% filter(str_detect(term, 'prot')) %>% select(estimate))
ar_alphai_log <- mt_loght %*% as.matrix(rs_hgt_prot_log_tidy %>% filter(str_detect(term, 'prot')) %>% select(estimate))

# Child Weight
ar_beta <- rep(1/length(ar_Ai_lin), times=length(ar_Ai_lin))

# Estimation Results lin and log both store
mt_opti <- cbind(ar_alphai_lin, ar_alphai_log, ar_Ai_lin, ar_Ai_log, ar_beta)
ar_st_varnames <- c('alpha_lin', 'alpha_log', 'A_lin', 'A_log', 'beta')
df_esti_alpha_A_beta <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))

# estimation results file to export (keeping data keys)
df_esti <- bind_cols(df_hw_cebu_m24, df_esti_alpha_A_beta) %>%
              select(one_of(c('S.country', 'vil.id', 'indi.id', 'svymthRound', ar_st_varnames)))

# Initate Dataframe that will store all estimates and optimal allocation relevant information
mt_opti <- cbind(ar_alphai_lin, ar_Ai_lin, ar_beta)
ar_st_varnames <- c('alpha', 'A', 'beta')
tb_opti <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))

# Unique beta, A, and alpha groups
tb_opti_unique <- tb_opti %>% group_by(!!!syms(ar_st_varnames)) %>%
                    arrange(!!!syms(ar_st_varnames)) %>%
                    summarise(n_obs_group=n())
```

## Export Data to */data* Folder

I would like to be able to directly call the dataset generated here in various functions. Save the datafile we just created in the project folder.

```{r}
if (bl_save_rda) {
  df_opt_dtgch_cbem4 <- df_esti
  usethis::use_data(df_opt_dtgch_cbem4, df_opt_dtgch_cbem4, overwrite = TRUE)
}
```
