---
title: BINARY--Binary Provision Optimal Allocation Solution (Line by Line)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BINARY--Binary Provision Optimal Allocation Solution (Line by Line)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Back to **[Fan](https://fanwangecon.github.io/)**'s Optimal Allocation Homepage **[Table of Content](https://fanwangecon.github.io/PrjOptiAllo==c/)**

# Objective

The objective of this file is to solve the linear $N_i\in{0,1}$ and $H_i$ problem without invoking any functions.

File name *ffv_opt_sobin_rkone_allrw*:

- *opt*: optimal allocation project
- *sobin*: binary provision solution
- *rkone*: rank at one
- *allrw*: all code line by line raw original file

## Algorithm Outline

Given $N$ individuals, if each individual could or could not receive the provision, given finite resource, and common cost, if total resource avaible could finance input for $M$ individuals, there would be $\frac{N!}{(M-N)!\cdot N!)}$ number of possible choices. Even with 10 individuals,

1. Solve unconstrained relative optimal allocation solutions from the continuous optimal allocation problem.
2. Evaluate relative allocations when allocation for $q$ is equal to $1$
3. This generates a new function where the y-axis values show outcomes for each other individual when individual $q$ allocation is at $1$.
4. Treat the $m$ components of the function as slope and intercept, the $q$ remaining component is the $x$, for each individual, there is a unique $x$
5. The randk order fully determin the sequence of provisions. This is similar to how rank order is determined in the continuous problem, with just a tiny difference of $\alpha$ added in the numerator.

## Input and Output

There is a dataset with child attributes, nutritional inputs and outputs. Run regression to estimate some input output relationship first. Then generate required inputs for code.

1. Required Input
  + @param df tibble data table including variables using svr names below each row is potentially an individual who will receive alternative allocations
  + @param svr_A_i string name of the A_i variable, dot product of covariates and coefficients
  + @param svr_alpha_i string name of the alpha_i variable, individual specific elasticity information
  + @param svr_beta_i string name of the beta_i variable, relative preference weight for each child
  + @param svr_N_i string name of the vector of existing inputs, based on which to compute aggregate resource
  + @param fl_N_hat float total resource avaible for allocation, if not specific, sum by svr_N_i
  + @param fl_rho float preference for equality for the planner
  + @return a dataframe that expands the df inputs with additional results.
2. The structure assumes some regression has already taken place to generate the i specific variables listed. and

Doing this allows for lagged intereaction that are time specific in an arbitrary way.

# Load Packages and Data

## Load Dependencies

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)
library(knitr)
library(kableExtra)
```

## Load Data

Generate four categories by initial height and mother's education levels combinations.

```{r Load Packages and Process Data}
# Load Data
tb_mtcars <- as_tibble(rownames_to_column(mtcars, var = "carname")) %>% rowid_to_column()
attach(tb_mtcars)

# Summarize
str(tb_mtcars)
summary(tb_mtcars)

# Generate Discrete Version of continuous variables
tb_mtcars <- tb_mtcars %>%
    mutate(wgtLowHigh = cut(wt,
                             breaks=c(-Inf, 3.1, Inf),
                             labels=c("LowWgt","HighWgt"))) %>%
    mutate(dratLowHigh = cut(drat,
                             breaks=c(-Inf, 3.59, Inf),
                             labels=c("lowRearAxleRatio","highRearAxleRatio"))) 

# Relabel some factors
tb_mtcars$vs <- factor(tb_mtcars$vs,
                        levels = c(0,1),
                        labels = c("engineVShaped", "engineStraight"))
tb_mtcars$am <- factor(tb_mtcars$am,
                        levels = c(0,1),
                        labels = c("automatic", "manual"))

# tabulate
tb_mtcars %>%
  group_by(am, vs) %>%
  summarize(freq = n()) %>% 
  pivot_wider(names_from = vs, values_from = freq)
```

# Testing the Equation

## Fraction Principle Testing

i value is larger than j value

```{r Formula Parameter Testing Basic}

alpha = 10
A_j_l = 0.4
A_i_l = 0.3

A_j_h = 1.4
A_i_h = 1.3

for (it_set in 1:4) {
  if(it_set == 1 || it_set == 2) {
    rho = 0.75
    if (it_set==1){
      A_j = A_j_l
      A_i = A_i_l
    } else if(it_set == 2) {
      A_j = A_j_h
      A_i = A_i_h
    }
  } else if(it_set == 3 || it_set == 4) {
    rho = -1
    if (it_set==3){
      A_j = A_j_l
      A_i = A_i_l
    } else if(it_set == 4) {
      A_j = A_j_h
      A_i = A_i_h
    }  
  }
  fl_num = (A_i+alpha)^rho - (A_j+alpha)^rho
  fl_den = (A_i)^rho - (A_j)^rho
  fl_rat = fl_num/fl_den
  cat('it_set:', it_set, ', fl_num:', fl_num, ', fl_den:', fl_den, ', fl_rat:', fl_rat, '\n')
}
```

## Testing Formula
```{r Formula Parameter Testing Basic}
alpha = 1
A_j_l = 0.3
A_i_l = 0.4
A_j_h = 1.3
A_i_h = 1.4
for (it_set in 1:4) {
  if(it_set == 1 || it_set == 2) {
    rho = 0.75
    if (it_set==1){
      A_j = A_j_l
      A_i = A_i_l
    } else if(it_set == 2) {
      A_j = A_j_h
      A_i = A_i_h
    }
  } else if(it_set == 3 || it_set == 4) {
    rho = -1
    if (it_set==3){
      A_j = A_j_l
      A_i = A_i_l
    } else if(it_set == 4) {
      A_j = A_j_h
      A_i = A_i_h
    }  
  }
  fl_num = (A_j+alpha)^rho - (A_j)^rho
  fl_den = (A_i+alpha)^rho - (A_i)^rho
  fl_rat = fl_num/fl_den
  cat('it_set:', it_set, ', fl_num:', fl_num, ', fl_den:', fl_den, ', fl_rat:', fl_rat, '\n')
}
```
```{r Formula Parameter Testing Common Params}
alpha = 0.1
A_j_l = 0.4
A_i_l = 0.5
A_j_h = 1.4
A_i_h = 1.5
for (it_set in 1:4) {
  if(it_set == 1 || it_set == 2) {
    rho = 0.75
    if (it_set==1){
      A_j = A_j_l
      A_i = A_i_l
    } else if(it_set == 2) {
      A_j = A_j_h
      A_i = A_i_h
    }
  } else if(it_set == 3 || it_set == 4) {
    alpha = 0.5
    rho = -10
    if (it_set==3){
      A_j = A_j_l
      A_i = A_i_l
    } else if(it_set == 4) {
      A_j = A_j_h
      A_i = A_i_h
    }  
  }
  fl_num = (A_j+alpha)^rho - (A_i+alpha)^rho
  fl_den = (A_j)^rho - (A_i)^rho
  fl_rat = fl_num/fl_den
  cat('it_set:', it_set, ', fl_num:', fl_num, ', fl_den:', fl_den, ', fl_rat:', fl_rat, '\n')
}
```

# Regression with Data and Construct Input Arrays

## Regress MPG on Manual Transmission

```{r Linear Regression Testing}
# A. Regree MPG on horse power and binary if Manual or not (manual = 1)
# 1. more horse power less MPG
# 2. manual larger MPG
print(summary(lm(mpg ~ hp + wt + factor(am) - 1)))

# B. Also incorporate now engine shape, vs = 0 if v-shaped
print(summary(lm(mpg ~ hp + factor(am) - 1)))

print(summary(lm(mpg ~ hp + carb + factor(am):factor(vs) - 1)))

#
print(summary(lm(mpg ~ hp + carb + factor(vs) + factor(am):factor(vs))))

# Preferred way of appearing, in the following regression we can se:
# 1. more more power less MPG, Straight engine slightly higher MPG
# 2. V-shape engine car, going from auto to manual trans gain 4.1 MPG
# 2. straight shape engine, going from auto to manual trans gain 6.67 MPG

# Store Regression Results
mt_model <- model.matrix(~ hp + qsec + factor(vs) + factor(am):factor(vs))
rs_mpg_on_auto = lm(mpg ~ mt_model - 1)
print(summary(rs_mpg_on_auto))
rs_mpg_on_auto_tidy = tidy(rs_mpg_on_auto)
rs_mpg_on_auto_tidy
```

## Construct Input Arrays $A_i$ and $\alpha_i$

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Post Regression Input Processing}
# Estimates Table
head(rs_mpg_on_auto_tidy, 6)
# Covariates
head(mt_model, 5)

# Covariates coefficients from regression (including constant)
ar_fl_cova_esti <- as.matrix(rs_mpg_on_auto_tidy %>% filter(!str_detect(term, 'am')) %>% select(estimate))
ar_fl_main_esti <- as.matrix(rs_mpg_on_auto_tidy %>% filter(str_detect(term, 'am')) %>% select(estimate))
head(ar_fl_cova_esti, 5)
head(ar_fl_main_esti, 5)

# Select Matrix subcomponents
mt_cova <- as.matrix(as_tibble(mt_model) %>% select(-contains("am")))
mt_intr <- model.matrix(~ factor(am) - 1)

# Generate A_i, use mt_cova_wth_const
ar_A_m <- mt_cova %*% ar_fl_cova_esti
head(ar_A_m, 5)

# Generate alpha_i
ar_alpha_m <- mt_intr %*% ar_fl_main_esti
head(ar_alpha_m, 5)
```

## Individual Weight

```{r}
# Child Weight
ar_beta_m <- rep(1/length(ar_A_m), times=length(ar_A_m))
```

## Matrix with Required Inputs for Allocation 

```{r}
# Initate Dataframe that will store all estimates and optimal allocation relevant information
# combine identifying key information along with estimation A, alpha results
# note that we only need indi.id as key
mt_opti <- cbind(ar_alpha_m, ar_A_m, ar_beta_m)
ar_st_varnames <- c('alpha', 'A', 'beta')
df_esti_alpha_A_beta <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))
tb_key_alpha_A_beta <- bind_cols(tb_mtcars, df_esti_alpha_A_beta) %>%
              select(one_of(c('rowid', 'carname', 'mpg', 'hp', 'qsec', 'vs', 'am', ar_st_varnames)))

# Unique beta, A, and alpha check
tb_opti_unique <- tb_key_alpha_A_beta %>% group_by(!!!syms(ar_st_varnames)) %>%
                    arrange(!!!syms(ar_st_varnames)) %>%
                    summarise(n_obs_group=n())

# Show cars
head(tb_key_alpha_A_beta, 32)
```

# Optimal Linear Allocations

## Parameters for Optimal Allocation

```{r Set Allocation Parameters}
# Child Count
it_obs = dim(tb_opti_unique)[1]

# Vector of Planner Preference
ar_rho <- c(-100, 0.8)
ar_rho <- c(-50, -25, -10)
ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.8)
ar_rho <- c(seq(-200, -100, length.out=5), seq(-100, -25, length.out=5), seq(-25, -5, length.out=5), seq(-5, -1, length.out=5), seq(-1, -0.01, length.out=5), seq(0.01, 0.25, length.out=5), seq(0.25, 0.90, length.out=5))
ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.99)
ar_rho <- c(-100, -10, -1, 0.05, 0.9)
ar_rho <- unique(ar_rho)
```

## Optimal binary Allocation (CRS)

This also works with any CRS CES.

```{r Optimal Linear Allocation Hard Code All Rho}
# Optimal Linear Equation

# Pull arrays out
ar_A <- tb_key_alpha_A_beta %>% pull(A)
ar_alpha <- tb_key_alpha_A_beta %>% pull(alpha)
ar_beta <- tb_key_alpha_A_beta %>% pull(beta)

# Define Function for each individual m, with hard coded arrays
ffi_binary_dplyrdo_func <- function(ls_row, fl_rho, bl_old=FALSE){
  # @param bl_old, weather to use old incorrect solution
  # hard coded inputs are:
  # 1, ar_A
  # 2, ar_alpha
  # 3, ar_beta
  # note follow https://fanwangecon.github.io/R4Econ/support/function/fs_applysapplymutate.html
  
  fl_alpha <- ls_row$alpha
  fl_A <- ls_row$A
  fl_beta <- ls_row$beta
  
  if (bl_old) {
    
    # the equation below is like doubly wrong I think, should be 1/(fl_rho-1)
    bl_old_wrong = TRUE
    if (bl_old_wrong) {
      ar_left <- ((ar_A + ar_alpha)/(fl_A + fl_alpha))
      ar_right <- ((ar_alpha*ar_beta)/(fl_alpha*fl_beta))^(1/(1-fl_rho))
      ar_full <- ar_left*ar_right
      ar_indicator <- (ar_full <= 1)
    } else {
      ar_left <- ((ar_A + ar_alpha)/(fl_A + fl_alpha))
      ar_right <- ((ar_alpha*ar_beta)/(fl_alpha*fl_beta))^(1/(fl_rho-1))
      ar_full <- ar_left*ar_right
      ar_indicator <- (ar_full <= 1)
    }
    
  } else {
      ar_left <- (
                  ((ar_A + ar_alpha)^fl_rho - (ar_A)^fl_rho)
                  /
                  ((fl_A + fl_alpha)^fl_rho - (fl_A)^fl_rho)
                 )
      ar_right <- ((ar_beta)/(fl_beta))
      ar_full <- ar_left*ar_right
      ar_indicator <- (ar_full >= 1)
  }
  
  it_rank <- sum(ar_indicator)
  return(it_rank)
}
ffi_binary_dplyrdo_func(tb_key_alpha_A_beta[1,], 0)


# accumulate allocation results
tb_opti_alloc_all_rho <- tb_key_alpha_A_beta

# A. First Loop over Planner Preference
# Generate Rank Order
for (it_rho_ctr in seq(1,length(ar_rho))) {
  rho = ar_rho[it_rho_ctr]

  queue_rank <- tb_key_alpha_A_beta %>% rowwise() %>%
                              do(rk = ffi_binary_dplyrdo_func(., rho)) %>%
                              unnest(rk) %>% pull(rk)
  
  tb_with_rank <- tb_key_alpha_A_beta %>% add_column(queue_rank)

  # m. Keep for df collection individual key + optimal allocation
  # _on stands for optimal nutritional choices
  # _eh stands for expected height
  tb_opti_allocate_wth_key <- tb_with_rank %>% select(one_of('rowid','queue_rank')) %>%
                                rename(!!paste0('rho_c', it_rho_ctr, '_rk') := !!sym('queue_rank'))

  # n. merge optimal allocaiton results from different planner preference
  tb_opti_alloc_all_rho <- tb_opti_alloc_all_rho %>% left_join(tb_opti_allocate_wth_key, by='rowid')
}

# o. print results
print(summary(tb_opti_alloc_all_rho))
str(tb_opti_alloc_all_rho)

# Make Longer
st_bisec_prefix <- 'rho_c'
svr_abfafb_long_name <- 'rho'
svr_bisect_iter <- 'nothing'
svr_number_col <- 'rank'
tb_opti_alloc_all_rho_long <- tb_opti_alloc_all_rho %>%
  pivot_longer(
    cols = starts_with(st_bisec_prefix),
    names_to = c(svr_abfafb_long_name, svr_bisect_iter),
    names_pattern = paste0(st_bisec_prefix, "(.*)_(.*)"),
    values_to = svr_number_col
  )
```

### Bump Plot for Optimal Binary Allocations

```{r custom theme online}
# custom theme by Dominik Koch
my_theme <- function() {
  
  # Colors
  color.background = "white"
  color.text = "#22211d"
  
  # Begin construction of chart
  theme_bw(base_size=15) +
    
    # Format background colors
    theme(panel.background = element_rect(fill=color.background, color=color.background)) +
    theme(plot.background  = element_rect(fill=color.background, color=color.background)) +
    theme(panel.border     = element_rect(color=color.background)) +
    theme(strip.background = element_rect(fill=color.background, color=color.background)) +
    
    # Format the grid
    theme(panel.grid.major.y = element_blank()) +
    theme(panel.grid.minor.y = element_blank()) +
    theme(axis.ticks       = element_blank()) +
    
    # Format the legend
    theme(legend.position = "none") +
    
    # Format title and axis labels
    theme(plot.title       = element_text(color=color.text, size=20, face = "bold")) +
    theme(axis.title.x     = element_text(size=14, color="black", face = "bold")) +
    theme(axis.title.y     = element_text(size=14, color="black", face = "bold", vjust=1.25)) +
    theme(axis.text.x      = element_text(size=10, vjust=0.5, hjust=0.5, color = color.text)) +
    theme(axis.text.y      = element_text(size=10, color = color.text)) +
    theme(strip.text       = element_text(face = "bold")) +
    
    # Plot margins
    theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}
```

```{r binary allocation bump plot, fig.height = 10, fig.width = 10, fig.align = "center"}

tb_opti_alloc_all_rho_long %>%
  ggplot(aes(x = rho, y = rank, group = carname)) +
    geom_line(aes(color = carname, alpha = 1), size = 2) +
    geom_point(aes(color = carname, alpha = 1), size = 4) +
    scale_x_discrete(expand = c(0.85,0))+
    scale_y_reverse(breaks = 1:nrow(tb_opti_alloc_all_rho_long))+
    theme(legend.position = "none") +
    labs(x = "Equality vs Efficiency",
         y = "Rank",
         title = "Binary Allocation Rank") +
    my_theme() + 
    geom_text(data =tb_opti_alloc_all_rho,aes(y=rho_c1_rk,x=0.6,label=carname),hjust="right")

```

# Linear optimal Allocation Statistics and Graphs

## Compute Gini Statistics for Allocations and Outcomes

Compute Gini coefficient for each expected outcome and input vector, and store results in separate vectors for input allocation and expected outcomes, then combine.

See how **apply** works here: [R use Apply, Sapply and dplyr Mutate to Evaluate one Function Across Rows of a Matrix](https://fanwangecon.github.io/R4Econ/support/function/fs_applysapplymutate.html) from [R4Econ](https://fanwangecon.github.io/R4Econ).

See how gini function works: [ff_dist_gini_vector_pos](https://fanwangecon.github.io/REconTools/reference/ff_dist_gini_vector_pos.html) from [REconTools](https://fanwangecon.github.io/REconTools).

```{r Compute Gini}
# Select out columns for _on = optimal nutrition
mt_opti_alloc_all_rho <- data.matrix(tb_opti_alloc_all_rho %>% select(ends_with('_on')))
mt_expc_outcm_all_rho <- data.matrix(tb_opti_alloc_all_rho %>% select(ends_with('_eh')))

# Compute gini for each rho
ar_opti_alloc_gini <- suppressMessages(apply(t(mt_opti_alloc_all_rho), 1, ff_dist_gini_vector_pos))
ar_expc_outcm_gini <- suppressMessages(apply(t(mt_expc_outcm_all_rho), 1, ff_dist_gini_vector_pos))

# Combine results
# column names look like: rho_c1_on rho_c2_on rho_c3_on rho_c1_eh rho_c2_eh rho_c3_eh
tb_gini_onerow_wide <- cbind(as_tibble(t(ar_opti_alloc_gini)), as_tibble(t(ar_expc_outcm_gini)))
tb_gini_long <- tb_gini_onerow_wide %>%
  pivot_longer(
    cols = starts_with("rho"),
    names_to = c("it_rho_ctr", "oneh"),
    names_pattern = "rho_c(.*)_(.*)",
    values_to = "gini"
  ) %>%
  pivot_wider(
    id_cols = it_rho_ctr,
    names_from = oneh,
    values_from = gini
  )
planner_elas <- log(1/(1-ar_rho)+2)
tb_gini_long <- cbind(tb_gini_long, planner_elas)
```


### Gini Graphs

```{r Gini Allocation and Outcome Scatter Correlation}

scatter <- tb_gini_long %>%
      ggplot(aes(x=on, y=eh)) +
      geom_point(size=planner_elas) +
      labs(title = 'Scatter Plot of Optimal Allocation and Expected Outcomes Gini',
           x = paste0('GINI for Optimal Allocations'),
           y = paste0('GINI for Exp Outcome given Optimal Allocations'),
           caption = 'Optimal Allocations and Expected Outcomes, Cebu Height and Protein Example (Wang 2020)') +
      theme_bw()
print(scatter)
```


## Graphs Based on Individual Allocations
```{r Individal Allocation Plots}
# p. graph scatter of opposing preference correlation
svr_lowrho_ctr <- 1
svr_higrho_ctr <- length(ar_rho)-1
svr_lowrho <- paste0('rho_c', svr_lowrho_ctr, '_on')
svr_higrho <- paste0('rho_c', svr_higrho_ctr, '_eh')
tb_min_max_rho <- tb_opti_alloc_all_rho %>% select(one_of(c(svr_lowrho, svr_higrho)))

scatter <- tb_min_max_rho %>%
      ggplot(aes(x=!!sym(svr_lowrho), y=!!sym(svr_higrho))) +
      geom_point(size=1) +
      labs(title = 'Scatter Plot of Optimal Allocation Low and High Rho',
           x = paste0('Allocation High Rho = ', ar_rho[svr_higrho_ctr], ', More Efficient'),
           y = paste0('Allocation Low Rho = ', ar_rho[svr_lowrho_ctr], ', More Equal'),
           caption = 'Optimal linear allocation results, given height reg on protein results. (Wang 2020)') +
      theme_bw()
print(scatter)

# lineplot <- tb_opti %>%
#     gather(variable, value, -month) %>%
#     ggplot(aes(x=month, y=value, colour=variable, linetype=variable)) +
#         geom_line() +
#         geom_point() +
#         labs(title = 'Mean and SD of Temperature Acorss US Cities',
#              x = 'Months',
#              y = 'Temperature in Fahrenheit',
#              caption = 'Temperature data 2017') +
#         scale_x_continuous(labels = as.character(df_temp_mth_summ$month),
#                            breaks = df_temp_mth_summ$month)

```

