---
title: DISCRETE--Discrete Provision Optimal Allocation Solution (Line by Line)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BINARY--Binary Provision Optimal Allocation Solution (Line by Line)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Back to **[Fan](https://fanwangecon.github.io/)**'s Optimal Allocation Homepage **[Table of Content](https://fanwangecon.github.io/PrjOptiAllo==c/)**

# Objective

The effects of mother's smoking status on birth weight. If we aim to improve health outcomes, what is the priority sequence of mothers whom we would want to discourage from smoking.

# Load Packages and Data

## Load Dependencies

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)

library(PrjOptiAlloc)

library(knitr)
library(kableExtra)
```

## Load Data

Generate four categories by initial height and mother's education levels combinations.

```{r Load Packages and Process Data}
# Load Data
data(df_opt_birthwt)

# Summarize
str(df_opt_birthwt)
summary(df_opt_birthwt)

# Generate Discrete Version of continuous variables
df_opt_birthwt <- df_opt_birthwt %>%
    mutate(momwgtLowHigh = cut(lwt,
                               breaks=c(-Inf, 129, Inf),
                               labels=c("LW","HW"))) %>%
    mutate(mombirthage = cut(age,
                               breaks=c(-Inf, 24, Inf),
                               labels=c("young","older"))) %>%
    mutate(ftvm3 = cut(ftv,
                               breaks=c(-Inf, 1, 2, Inf),
                               labels=c("novisit","1visit","morevisit")))

attach(df_opt_birthwt)
summary(df_opt_birthwt)

# tabulate
df_opt_birthwt %>%
  group_by(race, smoke, ftv) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = race, values_from = freq)

summary(lm(bwt ~ lwt + factor(mombirthage) + factor(race) + factor(ht) + factor(ui)))
summary(lm(bwt ~ lwt + age + factor(race) + factor(race):factor(smoke) - 1))
summary(lm(bwt ~ lwt + age + factor(race) + factor(race):factor(smoke) - 1))

summary(lm(bwt ~ factor(ftv) - 1))
summary(lm(bwt ~ factor(ftv) + factor(ht) + factor(ui) + factor(ptl) - 1))

summary(lm(bwt ~ lwt + factor(race) + (ftv) - 1))
```

# Regression with Data and Construct Input Arrays

## Binary Problem Regress Birth Weight on Mother's Smoking Status

```{r Linear Regression Testing}
# A. Regress child birth weight on mother's age, race and mother's weight
print(summary(lm(bwt ~ lwt + factor(mombirthage) + factor(race))))
print(summary(lm(bwt ~ lwt + factor(mombirthage) + factor(race) + factor(ht) + factor(ui))))
print(summary(lm(bwt ~ lwt + age + factor(race) + factor(ht) + factor(ui) + factor(ptl))))

print(summary(lm(bwt ~ lwt + age + factor(race) + factor(smoke))))
print(summary(lm(bwt ~ lwt + age + factor(race) + factor(ht) + factor(ui) +  factor(smoke))))

print(summary(lm(bwt ~ lwt + age + factor(momwgtLowHigh):factor(smoke) - 1)))
print(summary(lm(bwt ~ lwt + age + factor(ht) + factor(ui) + factor(race) + factor(race):factor(smoke))))
print(summary(lm(bwt ~ lwt + age + factor(race) + factor(race):factor(smoke) - 1)))

# Preferred way of appearing, in the following regression we can se:
# 1. more more power less MPG, Straight engine slightly higher MPG
# 2. V-shape engine car, going from auto to manual trans gain 4.1 MPG
# 2. straight shape engine, going from auto to manual trans gain 6.67 MPG

# Store Regression Results
mt_model <- model.matrix(~ lwt + age + factor(ht) + factor(ui) + factor(ptl) + factor(race) + factor(smoke):factor(race))
rs_wgt_on_smke = lm(bwt ~ mt_model - 1)
print(summary(rs_wgt_on_smke))
rs_wgt_on_smke_tidy = tidy(rs_wgt_on_smke)
rs_wgt_on_smke_tidy
```

## Construct Input Arrays $A_i$ and $\alpha_i$

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Post Regression Input Processing}
# Estimates Table
head(rs_wgt_on_smke_tidy, 6)
# Covariates
head(mt_model, 5)

# Covariates coefficients from regression (including constant)
ar_fl_cova_esti <- as.matrix(rs_wgt_on_smke_tidy %>% filter(!str_detect(term, 'smoke')) %>% select(estimate))
ar_fl_main_esti <- as.matrix(rs_wgt_on_smke_tidy %>% filter(str_detect(term, 'smoke')) %>% select(estimate))
head(ar_fl_cova_esti, 5)
head(ar_fl_main_esti, 5)

# Select Matrix subcomponents
mt_cova <- as.matrix(as_tibble(mt_model) %>% select(-contains("smoke")))
mt_intr <- model.matrix(~ factor(race) - 1)

# Generate A_i, use mt_cova_wth_const
ar_A_m <- mt_cova %*% ar_fl_cova_esti
head(ar_A_m, 5)

# Generate alpha_i
ar_alpha_m <- mt_intr %*% ar_fl_main_esti
head(ar_alpha_m, 5)
```

## Individual Weight

```{r}
# Child Weight
ar_beta_m <- rep(1/length(ar_A_m), times=length(ar_A_m))
```

## Matrix with Required Inputs for Allocation

```{r}
# Initate Dataframe that will store all estimates and optimal allocation relevant information
# combine identifying key information along with estimation A, alpha results
# note that we only need indi.id as key
mt_opti <- cbind(ar_alpha_m, ar_A_m, ar_beta_m)
ar_st_varnames <- c('alpha', 'A', 'beta')
df_esti_alpha_A_beta <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))
tb_key_alpha_A_beta <- bind_cols(df_opt_birthwt, df_esti_alpha_A_beta) %>%
              select(one_of(c('Index', 'age', 'lwt', 'race', 'smoke', 'ptl', 'ht', 'ui', ar_st_varnames)))

# Need to only include the smokers here
tb_key_alpha_A_beta <- tb_key_alpha_A_beta %>% filter(smoke == 1)

# Unique beta, A, and alpha check
tb_opti_unique <- tb_key_alpha_A_beta %>% group_by(!!!syms(ar_st_varnames)) %>%
                    arrange(!!!syms(ar_st_varnames)) %>%
                    summarise(n_obs_group=n())

# Show cars
head(tb_key_alpha_A_beta, 32)
```

# Optimal Linear Allocations

## Parameters for Optimal Allocation

```{r Set Allocation Parameters}
# Child Count
it_obs = dim(tb_opti_unique)[1]

# Vector of Planner Preference
ar_rho <- c(-100, 0.8)
ar_rho <- c(-50, -25, -10)
ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.8)
ar_rho <- c(seq(-200, -100, length.out=5), seq(-100, -25, length.out=5), seq(-25, -5, length.out=5), seq(-5, -1, length.out=5), seq(-1, -0.01, length.out=5), seq(0.01, 0.25, length.out=5), seq(0.25, 0.90, length.out=5))
ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.99)
ar_rho <- c(-100, -10, -1, 0.05, 0.9)
ar_rho <- unique(ar_rho)
```

## Optimal binary Allocation (CRS)

This also works with any CRS CES.

```{r Optimal Linear Allocation Hard Code All Rho}
# Optimal Linear Equation

# Pull arrays out
ar_A <- tb_key_alpha_A_beta %>% pull(A)
ar_alpha <- tb_key_alpha_A_beta %>% pull(alpha)
ar_beta <- tb_key_alpha_A_beta %>% pull(beta)

# Define Function for each individual m, with hard coded arrays
ffi_binary_dplyrdo_func <- function(ls_row, fl_rho, bl_old=FALSE){
  # @param bl_old, weather to use old incorrect solution
  # hard coded inputs are:
  # 1, ar_A
  # 2, ar_alpha
  # 3, ar_beta
  # note follow https://fanwangecon.github.io/R4Econ/support/function/fs_applysapplymutate.html

  fl_alpha <- ls_row$alpha
  fl_A <- ls_row$A
  fl_beta <- ls_row$beta

  if (bl_old) {

    # the equation below is like doubly wrong I think, should be 1/(fl_rho-1)
    bl_old_wrong = TRUE
    if (bl_old_wrong) {
      ar_left <- ((ar_A + ar_alpha)/(fl_A + fl_alpha))
      ar_right <- ((ar_alpha*ar_beta)/(fl_alpha*fl_beta))^(1/(1-fl_rho))
      ar_full <- ar_left*ar_right
      ar_indicator <- (ar_full <= 1)
    } else {
      ar_left <- ((ar_A + ar_alpha)/(fl_A + fl_alpha))
      ar_right <- ((ar_alpha*ar_beta)/(fl_alpha*fl_beta))^(1/(fl_rho-1))
      ar_full <- ar_left*ar_right
      ar_indicator <- (ar_full <= 1)
    }

  } else {
      ar_left <- (
                  ((ar_A + ar_alpha)^fl_rho - (ar_A)^fl_rho)
                  /
                  ((fl_A + fl_alpha)^fl_rho - (fl_A)^fl_rho)
                 )
      ar_right <- ((ar_beta)/(fl_beta))
      ar_full <- ar_left*ar_right
      ar_indicator <- (ar_full >= 1)
  }

  it_rank <- sum(ar_indicator)
  return(it_rank)
}
ffi_binary_dplyrdo_func(tb_key_alpha_A_beta[1,], 0)


# accumulate allocation results
tb_opti_alloc_all_rho <- tb_key_alpha_A_beta

# A. First Loop over Planner Preference
# Generate Rank Order
for (it_rho_ctr in seq(1,length(ar_rho))) {
  rho = ar_rho[it_rho_ctr]

  queue_rank <- tb_key_alpha_A_beta %>% rowwise() %>%
                              do(rk = ffi_binary_dplyrdo_func(., rho)) %>%
                              unnest(rk) %>% pull(rk)

  tb_with_rank <- tb_key_alpha_A_beta %>% add_column(queue_rank)

  # m. Keep for df collection individual key + optimal allocation
  # _on stands for optimal nutritional choices
  # _eh stands for expected height
  tb_opti_allocate_wth_key <- tb_with_rank %>% select(one_of('Index','queue_rank')) %>%
                                rename(!!paste0('rho_c', it_rho_ctr, '_rk') := !!sym('queue_rank'))

  # n. merge optimal allocaiton results from different planner preference
  tb_opti_alloc_all_rho <- tb_opti_alloc_all_rho %>% left_join(tb_opti_allocate_wth_key, by='Index')
}

# o. print results
print(summary(tb_opti_alloc_all_rho))
str(tb_opti_alloc_all_rho)

# Make Longer
st_bisec_prefix <- 'rho_c'
svr_abfafb_long_name <- 'rho'
svr_bisect_iter <- 'nothing'
svr_number_col <- 'rank'
tb_opti_alloc_all_rho_long <- tb_opti_alloc_all_rho %>%
  pivot_longer(
    cols = starts_with(st_bisec_prefix),
    names_to = c(svr_abfafb_long_name, svr_bisect_iter),
    names_pattern = paste0(st_bisec_prefix, "(.*)_(.*)"),
    values_to = svr_number_col
  )
```

### Bump Plot for Optimal Binary Allocations

```{r binary allocation bump plot, fig.height = 10, fig.width = 10, fig.align = "center"}

tb_opti_alloc_all_rho_long %>%
  ggplot(aes(x = rho, y = rank, group = Index)) +
    geom_line(aes(color = race, alpha = 1), size = 2) +
    geom_point(aes(color = race, alpha = 1), size = 4) +
    scale_x_discrete(expand = c(0.85,0))+
    scale_y_reverse(breaks = 1:nrow(tb_opti_alloc_all_rho_long))+
    theme(legend.position = "none") +
    labs(x = "Equality vs Efficiency",
         y = "Rank",
         title = "Binary Deallocation Rank, Weight and Smokng") +
    ffy_opt_ghthm_dk() +
    geom_text(data =tb_opti_alloc_all_rho,aes(y=rho_c1_rk,x=0.6,label=Index),hjust="right")

```
