---
title: BINARY--NSW Training WAGE Optimal Allocation Solution
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BINARY--NSW Training WAGE Optimal Allocation Solution}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Back to **[Fan](https://fanwangecon.github.io/)**'s Optimal Allocation Homepage **[Table of Content](https://fanwangecon.github.io/PrjOptiAllo==c/)**

# Objective

Test binary allocation queue with Lalonde training dataset. There are 722 observations, 297 in the treatment group, 425 in the control group.

Following Lalonda, regressions are in terms of wage levels (not log wage)

# Load Packages and Data

## Load Dependencies

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)

library(PrjOptiAlloc)

library(knitr)
library(kableExtra)
```

## Load Data and Generate New Variables

Generate four categories by initial height and mother's education levels combinations.

```{r Load Packages and Process Data}
# Load Data
data(df_opt_lalonde_training)
dft <- df_opt_lalonde_training %>% mutate(id = X) %>%
           select(-X) %>%
           select(id, everything())
# Summarize
str(dft)
summary(dft)

# Generate Employment 0 or 1 status Variable, and non-zero wages variable
dft <- dft %>% mutate(emp78 =
                      case_when(re78 <= 0 ~ 0,
                                TRUE ~ 1)) %>%
               mutate(emp75 =
                      case_when(re75 <= 0 ~ 0,
                                TRUE ~ 1)) %>%
               mutate(emp74 =
                      case_when(re74 <= 0 ~ 0,
                                TRUE ~ 1))

# Generate combine black + hispanic status
# 0 = white, 1 = black, 2 = hispanics
dft <- dft %>%
    mutate(race =
             case_when(black == 1 ~ 1,
                       hisp == 1 ~ 2,
                       TRUE ~ 0))

dft <- dft %>%
    mutate(age_m2 =
             case_when(age <= 23 ~ 1,
                       age >  23~ 2)) %>%
    mutate(age_m3 =
             case_when(age <= 20 ~ 1,
                       age > 20 & age <= 26 ~ 2,
                       age > 26 ~ 3))



# filter(re78 != 0) %>%
# mutate(re78_log = log(re78))

# Exclude zeros
# when this is on, both linear and log linear results exclude wage = 0
# dft <- dft %>%
#     filter(re78 > 0)

# Generate Discrete Version of continuous variables
# dft <- dft %>%
#     mutate(momwgtLowHigh = cut(lwt,
#                                breaks=c(-Inf, 129, Inf),
#                                labels=c("LW","HW"))) %>%
#     mutate(mombirthage = cut(age,
#                                breaks=c(-Inf, 24, Inf),
#                                labels=c("young","older")))
```

# Regression with Data and Construct Input Arrays

## Tabulate and Summarize Averages

What is the average difference in wage between treatment and control, do they match what is reported in Lalonda (1986)?

```{r wage treatment effects}

# Summarize average for all variables grouping by treatment status
# re78 is significantly different
round(t(dft %>% group_by(trt) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)
round(t(dft %>% group_by(marr, age_m2) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)

# Summarize by finer sub groups: RACE
# big increase for black, but not for other group
round(t(dft %>% group_by(trt, race) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)

# Summarize by finer sub groups: MARRIAGE
# big increase for black, but not for other group
round(t(dft %>% group_by(trt, marr) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)

# Summarize by finer sub groups: AGE GROUPS
round(t(dft %>% group_by(trt, age_m2) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)

round(t(dft %>% group_by(trt, age_m3) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)

round(t(dft %>% group_by(trt, marr, age_m2) %>%
        summarise_if(is.numeric, funs(mean = mean), na.rm = TRUE)), digits=3)

```

## Tabulate

```{r tabulate groups}
# Tabulate groups, how many in each group, enough for group heterogeneity in effects?
dft %>%
  group_by(trt, marr) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = trt, values_from = freq)

# Tabulate groups, how many in each group, enough for group heterogeneity in effects?
dft %>%
  group_by(trt, age_m2, marr) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = trt, values_from = freq)
```

## Regression Testing

```{r Regression Testing}

# As noted in Lalonda (1986), functional form assumptions do not matter much
# Dummies, treatment effect average about 801 dollars
summary(lm(re78 ~ factor(age)
                  + factor(educ)
                  + factor(race)
                  + factor(marr)
                  + factor(nodeg)
                  + factor(trt) - 1,
                  data = dft))

# cts Controls and treatment effects about 761 dollars
summary(lm(re78 ~  age + I(age^2) +
                   educ + I(educ^2) +
                  + factor(race)
                  + factor(marr)
                  + factor(nodeg)
                  + factor(trt) - 1,
                  data = dft))

# Treatment interactions by marriage status, 476 unmarried, vs 2216 for married
summary(lm(re78 ~  age + I(age^2) +
                   educ + I(educ^2) +
                  + factor(race)
                  + factor(marr)
                  + factor(nodeg)
                  + factor(marr):factor(trt) - 1,
                  data = dft))

# Treatment interactions by marriage status, 453 vs 1070 each age group
summary(lm(re78 ~  age + I(age^2) + factor(age_m2) +
                   educ + I(educ^2) +
                  + factor(race)
                  + factor(marr)
                  + factor(nodeg)
                  + factor(age_m2):factor(trt) - 1,
                  data = dft))

# Treatment interactions by marriage status, greater effect for older married + younger married
summary(lm(re78 ~  age + I(age^2) + factor(age_m2) +
                   educ + I(educ^2) +
                  + factor(race)
                  + factor(marr)
                  + factor(nodeg)
                  + factor(marr):factor(age_m2):factor(trt) - 1,
                  data = dft))

```

## Regress Wage on Training Status

### Linear Binary Problem

```{r Binary Linear Regression}
# Store Regression Results
mt_model <- model.matrix( ~ age + I(age^2) + factor(age_m2) +
                            educ + I(educ^2) +
                          + factor(race)
                          + factor(marr)
                          + factor(nodeg)
                          + factor(age_m2):factor(trt),
                          data = dft)
rs_wage_on_trt = lm(re78 ~ mt_model - 1, data = dft)
print(summary(rs_wage_on_trt))
rs_wage_on_trt_tidy = tidy(rs_wage_on_trt)
rs_wage_on_trt_tidy
```

## Construct Input Arrays $A_i$ and $\alpha_i$

### Linear Binary Regression

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Linear Binary Post Regression Input Processing}
# Estimates Table
head(rs_wage_on_trt_tidy, 6)
# Covariates
head(mt_model, 5)

# Covariates coefficients from regression (including constant)
ar_fl_cova_esti <- as.matrix(rs_wage_on_trt_tidy %>% filter(!str_detect(term, 'trt')) %>% select(estimate))
ar_fl_main_esti <- as.matrix(rs_wage_on_trt_tidy %>% filter(str_detect(term, 'trt')) %>% select(estimate))
head(ar_fl_cova_esti, 5)
head(ar_fl_main_esti, 5)

# Select Matrix subcomponents
mt_cova <- as.matrix(as_tibble(mt_model) %>% select(-contains("trt")))
mt_intr <- model.matrix(~ factor(marr) - 1, data = dft)

# Generate A_i, use mt_cova_wth_const
ar_A_m <- mt_cova %*% ar_fl_cova_esti
head(ar_A_m, 5)

# Generate alpha_i
ar_alpha_m <- mt_intr %*% ar_fl_main_esti
head(ar_alpha_m, 5)
```

## Individual Weight
```{r}
# Child Weight
ar_beta_m <- rep(1/length(ar_A_m), times=length(ar_A_m))
```

## Matrix with Required Inputs for Allocation

### Linear Binary Matrix

```{r Linear Allocattion Space Matrix}
# Initate Dataframe that will store all estimates and optimal allocation relevant information
# combine identifying key information along with estimation A, alpha results
# note that we only need indi.id as key
mt_opti <- cbind(ar_alpha_m, ar_A_m, ar_beta_m)
ar_st_varnames <- c('alpha', 'A', 'beta')
df_esti_alpha_A_beta <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))
tb_key_alpha_A_beta <- bind_cols(dft, df_esti_alpha_A_beta) %>%
              select(one_of(c('id', 'trt', 'age', 'educ', 'race', 'marr', 'nodeg', 're78',
                              ar_st_varnames)))

# Rescale A and alpha to deal more easily with large powers
tb_key_alpha_A_beta <- tb_key_alpha_A_beta %>%
                          mutate(alpha = alpha/1000, A = A/1000)

# Need to only include the smokers here
# tb_key_alpha_A_beta <- tb_key_alpha_A_beta %>% filter(trt == 0)

# Unique beta, A, and alpha check
tb_opti_unique <- tb_key_alpha_A_beta %>% group_by(!!!syms(ar_st_varnames)) %>%
                    arrange(!!!syms(ar_st_varnames)) %>%
                    summarise(n_obs_group=n())

# Show cars
head(tb_key_alpha_A_beta, 32)
```

# Optimal Linear Allocations

```{r Set Allocation Parameters}
# Child Count
it_obs = dim(tb_opti_unique)[1]

# Vector of Planner Preference
ar_rho <- c(-100, 0.8)
ar_rho <- c(-50, -25, -10)
ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.8)
ar_rho <- c(seq(-200, -100, length.out=5), seq(-100, -25, length.out=5), seq(-25, -5, length.out=5), seq(-5, -1, length.out=5), seq(-1, -0.01, length.out=5), seq(0.01, 0.25, length.out=5), seq(0.25, 0.90, length.out=5))
# ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.99)
# ar_rho <- c(-20, -1, 0.05, 0.9)
# ar_rho <- c(-50, -40, -30, -20, -15, -10, -7.5, -5,-3,-2,-1)
# ar_rho = c(-100, -0.001,  0.95)
ar_rho = c(-100, -0.001,  0.95)
ar_rho <- unique(ar_rho)

ls_bin_solu_all_rhos <-
  ffp_opt_anlyz_rhgin_bin(tb_key_alpha_A_beta, svr_id_i = 'id',
                          svr_A_i = 'A', svr_alpha_i = 'alpha', svr_beta_i = 'beta',
                          ar_rho = ar_rho,
                          svr_inpalc = 'opti_alloc_queue',
                          svr_expout = 'opti_exp_outcome')

df_all_rho <- ls_bin_solu_all_rhos$df_all_rho
df_all_rho_long <- ls_bin_solu_all_rhos$df_all_rho_long

# How many people have different ranks across rhos
it_how_many_vary_rank <- sum(df_all_rho$rank_max - df_all_rho$rank_min)
it_how_many_vary_rank
```

## Save Results

```{r saving file}

# Change Variable names so that this can becombined with logit file later
df_all_rho <- df_all_rho %>% rename_at(vars(starts_with("rho_")), funs(str_replace(., "rk", "rk_wage")))
df_all_rho <- df_all_rho %>% rename(A_i_wage = A, alpha_i_wage = alpha, beta_i_wage = beta,
                                    rank_min_wage = rank_min, rank_max_wage = rank_max, avg_rank_wage = avg_rank)

# Save File
df_opt_lalonde_training_wage <- df_all_rho
usethis::use_data(df_opt_lalonde_training_wage, df_opt_lalonde_training_wage, overwrite = TRUE)
```

## Bump Plot for Optimal Binary Allocations

```{r binary allocation bump plot, fig.height = 10, fig.width = 20, fig.align = "center"}

# tb_opti_alloc_all_rho_long %>%
#   ggplot(aes(x = rho, y = rank, group = id)) +
#     geom_line(aes(color = race, alpha = 1), size = 2) +
#     geom_point(aes(color = race, alpha = 1), size = 4) +
#     scale_x_discrete(expand = c(0.85,0))+
#     scale_y_reverse(breaks = 1:nrow(tb_opti_alloc_all_rho_long))+
#     theme(legend.position = "none") +
#     labs(x = "Equality vs Efficiency",
#          y = "Rank",
#          title = "Binary Allocation Rank, which untrained to receive training first") +
#     ffy_opt_ghthm_dk() +
#     geom_text(data =tb_opti_alloc_all_rho,aes(y=rho_c1_rk,x=0.6,label=id),hjust="right")

```
