---
title: BINARY--Binary Provision Optimal Allocation Solution (Line by Line)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BINARY--Binary Provision Optimal Allocation Solution (Line by Line)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Back to **[Fan](https://fanwangecon.github.io/)**'s Optimal Allocation Homepage **[Table of Content](https://fanwangecon.github.io/PrjOptiAlloc/)**

# Objective

The objective of this file is to solve the linear $N_i\in{0,1}$ and $H_i$ problem without invoking any functions.

File name *ffv_opt_sobin_rkone_allrw*:

- *opt*: optimal allocation project
- *sobin*: binary provision solution
- *rkone*: rank at one
- *allrw*: all code line by line raw original file

## Algorithm Outline

Given $N$ individuals, if each individual could or could not receive the provision, given finite resource, and common cost, if total resource avaible could finance input for $M$ individuals, there would be $\frac{N!}{(M-N)!\cdot N!)}$ number of possible choices. Even with 10 individuals,

1. Solve unconstrained relative optimal allocation solutions from the continuous optimal allocation problem.
2. Evaluate relative allocations when allocation for $q$ is equal to $1$
3. This generates a new function where the y-axis values show outcomes for each other individual when individual $q$ allocation is at $1$.
4. Treat the $m$ components of the function as slope and intercept, the $q$ remaining component is the $x$, for each individual, there is a unique $x$
5. The randk order fully determin the sequence of provisions. This is similar to how rank order is determined in the continuous problem, with just a tiny difference of $\alpha$ added in the numerator.

## Input and Output

There is a dataset with child attributes, nutritional inputs and outputs. Run regression to estimate some input output relationship first. Then generate required inputs for code.

1. Required Input
  + @param df tibble data table including variables using svr names below each row is potentially an individual who will receive alternative allocations
  + @param svr_A_i string name of the A_i variable, dot product of covariates and coefficients
  + @param svr_alpha_i string name of the alpha_i variable, individual specific elasticity information
  + @param svr_beta_i string name of the beta_i variable, relative preference weight for each child
  + @param svr_N_i string name of the vector of existing inputs, based on which to compute aggregate resource
  + @param fl_N_hat float total resource avaible for allocation, if not specific, sum by svr_N_i
  + @param fl_rho float preference for equality for the planner
  + @return a dataframe that expands the df inputs with additional results.
2. The structure assumes some regression has already taken place to generate the i specific variables listed. and

Doing this allows for lagged intereaction that are time specific in an arbitrary way.

# Load Packages and Data

## Load Dependencies

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(stringr)
library(broom)
library(ggplot2)
library(REconTools)
library(knitr)
library(kableExtra)
```

## Load Data

Generate four categories by initial height and mother's education levels combinations.

```{r Load Packages and Process Data}
# Load Library

# Select Cebu Only
tb_mtcars <- mtcars

# Generate Discrete Version of momEdu
tb_mtcars <- tb_mtcars %>%
    mutate(wgtLowHigh = cut(wt,
                             breaks=c(-Inf, 3.2, Inf),
                             labels=c("LowWgt","HighWgt"))) %>%
    mutate(dratLowHigh = cut(drat,
                             breaks=c(-Inf, 3.59, Inf),
                             labels=c("lowRearAxleRatio","highRearAxleRatio")))

tb_mtcars$wgtLowHigh = as.factor(tb_mtcars$wgtLowHigh)
tb_mtcars$dratLowHigh = as.factor(tb_mtcars$dratLowHigh)

# Attach
attach(tb_mtcars)
```

# Regression with Data and Construct Input Arrays

## What is the E

We

```{r Linear Regression}
# Input Matrix
mt_lincv <- model.matrix(~ hp)
mt_linht <- model.matrix(~ wgtLowHigh:dratLowHigh - 1)

# Regress Height At Month 24 on Nutritional Inputs with controls
rs_mpg_on_auto = lm(mpg ~ am:gear - 1)
print(summary(rs_mpg_on_auto))
rs_mpg_on_auto_tidy = tidy(rs_mpg_on_auto)
```

## Log-Linear Regression

Now log-linear regressions, where inptut coefficient differs by input groups.

```{r Log-Linear Regression}
# Input Matrix Generation
mt_logcv <- model.matrix(~ hgt0 + wgt0)
mt_loght <- model.matrix(~ sex:hgt0med - 1)

# Log and log regression for month 24
rs_hgt_prot_log = lm(log(hgt) ~ log(prot):mt_loght + mt_logcv - 1)
print(summary(rs_hgt_prot_log))
rs_hgt_prot_log_tidy = tidy(rs_hgt_prot_log)
```

## Construct Input Arrays $A_i$ and $\alpha_i$

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Post Regression Input Processing}

# Generate A_i
ar_Ai_lin <- mt_lincv %*% as.matrix(rs_mpg_on_auto_tidy %>% filter(!str_detect(term, 'prot')) %>% select(estimate))
ar_Ai_log <- mt_logcv %*% as.matrix(rs_hgt_prot_log_tidy %>% filter(!str_detect(term, 'prot')) %>% select(estimate))

# Generate alpha_i
ar_alphai_lin <- mt_linht %*% as.matrix(rs_mpg_on_auto_tidy %>% filter(str_detect(term, 'prot')) %>% select(estimate))
ar_alphai_log <- mt_loght %*% as.matrix(rs_hgt_prot_log_tidy %>% filter(str_detect(term, 'prot')) %>% select(estimate))

# Child Weight
ar_beta <- rep(1/length(ar_Ai_lin), times=length(ar_Ai_lin))

# Initate Dataframe that will store all estimates and optimal allocation relevant information
# combine identifying key information along with estimation A, alpha results
# note that we only need indi.id as key
mt_opti <- cbind(ar_alphai_lin, ar_Ai_lin, ar_beta)
ar_st_varnames <- c('alpha', 'A', 'beta')
df_esti_alpha_A_beta <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))
tb_key_alpha_A_beta <- bind_cols(tb_mtcars, df_esti_alpha_A_beta) %>%
              select(one_of(c('S.country', 'vil.id', 'indi.id', 'svymthRound', ar_st_varnames)))

# Unique beta, A, and alpha groups
tb_opti_unique <- tb_key_alpha_A_beta %>% group_by(!!!syms(ar_st_varnames)) %>%
                    arrange(!!!syms(ar_st_varnames)) %>%
                    summarise(n_obs_group=n())
```

# Optimal Linear Allocations

## Parameters for Optimal Allocation

```{r Set Allocation Parameters}
# Child Count
tb_mtcars_full <- tb_mtcars
it_obs = dim(tb_mtcars)[1]

# Total Resource Count
ar_prot_data = tb_mtcars$prot
fl_N_agg = sum(ar_prot_data)

# Vector of Planner Preference
ar_rho <- c(-100, 0.8)
ar_rho <- c(-50, -25, -10)
ar_rho <- c(-100, -5, -1, 0.1, 0.6, 0.8)
ar_rho <- c(seq(-200, -100, length.out=5), seq(-100, -25, length.out=5), seq(-25, -5, length.out=5), seq(-5, -1, length.out=5), seq(-1, -0.01, length.out=5), seq(0.01, 0.25, length.out=5), seq(0.25, 0.90, length.out=5))
ar_rho <- unique(ar_rho)
```

## Optimal Linear Allocation (CRS)

This also works with any CRS CES.

```{r Optimal Linear Allocation Hard Code All Rho}
# Optimal Linear Equation

# accumulate allocation results
tb_opti_alloc_all_rho <- tb_key_alpha_A_beta

# A. First Loop over Planner Preference
# Generate Rank Order
for (it_rho_ctr in seq(1,length(ar_rho))) {
  rho = ar_rho[it_rho_ctr]

  # B. Generate V4, Rank Index Value, rho specific
  # tb_opti <- tb_opti %>% mutate(!!paste0('rv_', it_rho_ctr) := A/((alpha*beta))^(1/(1-rho)))
  tb_opti <- tb_key_alpha_A_beta %>% mutate(rank_val = A/((alpha*beta))^(1/(1-rho)))

  # c. Generate Rank Index
  tb_opti <- tb_opti %>% arrange(rank_val) %>% mutate(rank_idx = row_number())

  # d. Populate lowest index alpha, beta, and A to all rows
  tb_opti <- tb_opti %>% mutate(lowest_rank_A = A[rank_idx==1]) %>%
                mutate(lowest_rank_alpha = alpha[rank_idx==1]) %>%
                mutate(lowest_rank_beta = beta[rank_idx==1])

  # e. relative slope and relative intercept with respect to lowest index
  tb_opti <- tb_opti %>%
                mutate(rela_slope_to_lowest =
                         (((lowest_rank_alpha*lowest_rank_beta)/(alpha*beta))^(1/(rho-1))*(lowest_rank_alpha/alpha))
                      ) %>%
                mutate(rela_intercept_to_lowest =
                         ((((lowest_rank_alpha*lowest_rank_beta)/(alpha*beta))^(1/(rho-1))*(lowest_rank_A/alpha)) - (A/alpha))
                      )

  # f. cumulative sums
  tb_opti <- tb_opti %>%
                mutate(rela_slope_to_lowest_cumsum =
                         cumsum(rela_slope_to_lowest)
                      ) %>%
                mutate(rela_intercept_to_lowest_cumsum =
                         cumsum(rela_intercept_to_lowest)
                      )

  # g. inverting cumulative slopes and intercepts
  tb_opti <- tb_opti %>%
                mutate(rela_slope_to_lowest_cumsum_invert =
                         (1/rela_slope_to_lowest_cumsum)
                      ) %>%
                mutate(rela_intercept_to_lowest_cumsum_invert =
                         ((-1)*(rela_intercept_to_lowest_cumsum)/(rela_slope_to_lowest_cumsum))
                      )

  # h. Relative x-intercept points
  tb_opti <- tb_opti %>%
                mutate(rela_x_intercept =
                         (-1)*(rela_intercept_to_lowest/rela_slope_to_lowest)
                      )

  # i. Inverted relative x-intercepts
  tb_opti <- tb_opti %>%
                mutate(opti_lowest_spline_knots =
                         (rela_intercept_to_lowest_cumsum + rela_slope_to_lowest_cumsum*rela_x_intercept)
                      )

  # j. Sort by order of receiving transfers/subsidies
  tb_opti <- tb_opti %>% arrange(rela_x_intercept)

  # k. Find position of subsidy
  tb_opti <- tb_opti %>% arrange(opti_lowest_spline_knots) %>%
                mutate(tot_devi = opti_lowest_spline_knots - fl_N_agg) %>%
                arrange((-1)*case_when(tot_devi < 0 ~ tot_devi)) %>%
                mutate(allocate_lowest =
                         case_when(row_number() == 1 ~
                                     rela_intercept_to_lowest_cumsum_invert +
                                     rela_slope_to_lowest_cumsum_invert*fl_N_agg)) %>%
                mutate(allocate_lowest = allocate_lowest[row_number() == 1]) %>%
                mutate(opti_allocate =
                         rela_intercept_to_lowest +
                         rela_slope_to_lowest*allocate_lowest) %>%
                mutate(opti_allocate =
                         case_when(opti_allocate >= 0 ~ opti_allocate, TRUE ~ 0)) %>%
                mutate(allocate_total = sum(opti_allocate, na.rm=TRUE))

  # l. Predictes Expected choice: A + alpha*opti_allocate
  tb_opti <- tb_opti %>% mutate(opti_exp_outcome = A + alpha*opti_allocate)  

  # m. Keep for df collection individual key + optimal allocation
  # _on stands for optimal nutritional choices
  # _eh stands for expected height
  tb_opti_main_results <- tb_opti %>%
    select(-one_of(c('lowest_rank_alpha', 'lowest_rank_beta')))
  tb_opti_allocate_wth_key <- tb_opti %>% select(one_of('indi.id','opti_allocate', 'opti_exp_outcome')) %>%
                                rename(!!paste0('rho_c', it_rho_ctr, '_on') := !!sym('opti_allocate'),
                                       !!paste0('rho_c', it_rho_ctr, '_eh') := !!sym('opti_exp_outcome'))

  # n. merge optimal allocaiton results from different planner preference
  tb_opti_alloc_all_rho <- tb_opti_alloc_all_rho %>% left_join(tb_opti_allocate_wth_key, by='indi.id')

  # print subset
  if (it_rho_ctr == 1 | it_rho_ctr == length(ar_rho) | it_rho_ctr == round(length(ar_rho)/2)) {
    print('')
    print('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')
    print(paste0('xxx rho:', rho))
    print('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')
    print('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')

    print(summary(tb_opti_main_results))
  }
}

# o. print results
print(summary(tb_opti_alloc_all_rho))
str(tb_opti_alloc_all_rho)
```

# Linear optimal Allocation Statistics and Graphs

## Compute Gini Statistics for Allocations and Outcomes

Compute Gini coefficient for each expected outcome and input vector, and store results in separate vectors for input allocation and expected outcomes, then combine.

See how **apply** works here: [R use Apply, Sapply and dplyr Mutate to Evaluate one Function Across Rows of a Matrix](https://fanwangecon.github.io/R4Econ/support/function/fs_applysapplymutate.html) from [R4Econ](https://fanwangecon.github.io/R4Econ).

See how gini function works: [ff_dist_gini_vector_pos](https://fanwangecon.github.io/REconTools/reference/ff_dist_gini_vector_pos.html) from [REconTools](https://fanwangecon.github.io/REconTools).

```{r Compute Gini}
# Select out columns for _on = optimal nutrition
mt_opti_alloc_all_rho <- data.matrix(tb_opti_alloc_all_rho %>% select(ends_with('_on')))
mt_expc_outcm_all_rho <- data.matrix(tb_opti_alloc_all_rho %>% select(ends_with('_eh')))

# Compute gini for each rho
ar_opti_alloc_gini <- suppressMessages(apply(t(mt_opti_alloc_all_rho), 1, ff_dist_gini_vector_pos))
ar_expc_outcm_gini <- suppressMessages(apply(t(mt_expc_outcm_all_rho), 1, ff_dist_gini_vector_pos))

# Combine results
# column names look like: rho_c1_on rho_c2_on rho_c3_on rho_c1_eh rho_c2_eh rho_c3_eh
tb_gini_onerow_wide <- cbind(as_tibble(t(ar_opti_alloc_gini)), as_tibble(t(ar_expc_outcm_gini)))
tb_gini_long <- tb_gini_onerow_wide %>%
  pivot_longer(
    cols = starts_with("rho"),
    names_to = c("it_rho_ctr", "oneh"),
    names_pattern = "rho_c(.*)_(.*)",
    values_to = "gini"
  ) %>%
  pivot_wider(
    id_cols = it_rho_ctr,
    names_from = oneh,
    values_from = gini
  )
planner_elas <- log(1/(1-ar_rho)+2)
tb_gini_long <- cbind(tb_gini_long, planner_elas)
```


### Gini Graphs

```{r Gini Allocation and Outcome Scatter Correlation}

scatter <- tb_gini_long %>%
      ggplot(aes(x=on, y=eh)) +
      geom_point(size=planner_elas) +
      labs(title = 'Scatter Plot of Optimal Allocation and Expected Outcomes Gini',
           x = paste0('GINI for Optimal Allocations'),
           y = paste0('GINI for Exp Outcome given Optimal Allocations'),
           caption = 'Optimal Allocations and Expected Outcomes, Cebu Height and Protein Example (Wang 2020)') +
      theme_bw()
print(scatter)
```


## Graphs Based on Individual Allocations
```{r Individal Allocation Plots}
# p. graph scatter of opposing preference correlation
svr_lowrho_ctr <- 1
svr_higrho_ctr <- length(ar_rho)-1
svr_lowrho <- paste0('rho_c', svr_lowrho_ctr, '_on')
svr_higrho <- paste0('rho_c', svr_higrho_ctr, '_eh')
tb_min_max_rho <- tb_opti_alloc_all_rho %>% select(one_of(c(svr_lowrho, svr_higrho)))

scatter <- tb_min_max_rho %>%
      ggplot(aes(x=!!sym(svr_lowrho), y=!!sym(svr_higrho))) +
      geom_point(size=1) +
      labs(title = 'Scatter Plot of Optimal Allocation Low and High Rho',
           x = paste0('Allocation High Rho = ', ar_rho[svr_higrho_ctr], ', More Efficient'),
           y = paste0('Allocation Low Rho = ', ar_rho[svr_lowrho_ctr], ', More Equal'),
           caption = 'Optimal linear allocation results, given height reg on protein results. (Wang 2020)') +
      theme_bw()
print(scatter)

# lineplot <- tb_opti %>%
#     gather(variable, value, -month) %>%
#     ggplot(aes(x=month, y=value, colour=variable, linetype=variable)) +
#         geom_line() +
#         geom_point() +
#         labs(title = 'Mean and SD of Temperature Acorss US Cities',
#              x = 'Months',
#              y = 'Temperature in Fahrenheit',
#              caption = 'Temperature data 2017') +
#         scale_x_continuous(labels = as.character(df_temp_mth_summ$month),
#                            breaks = df_temp_mth_summ$month)

```

