---
title: Linear (CRS) Optimal Allocation Solution (No Function)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear (CRS) Optimal Allocation Solution (No Function)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

Back to **[Fan](https://fanwangecon.github.io/)**'s Optimal Allocation Homepage **[Table of Content](https://fanwangecon.github.io/PrjOptiAlloc/)**

# Outline

There is a dataset with child attributes, nutritional inputs and outputs. Run regression to estimate some input output relationship first. Then generate required inputs for code.

1. Required Input
  + @param df tibble data table including variables using svr names below each row is potentially an individual who will receive alternative allocations
  + @param svr_A_i string name of the A_i variable, dot product of covariates and coefficients
  + @param svr_alpha_i string name of the alpha_i variable, individual specific elasticity information
  + @param svr_beta_i string name of the beta_i variable, relative preference weight for each child
  + @param svr_N_i string name of the vector of existing inputs, based on which to compute aggregate resource
  + @param fl_N_hat float total resource avaible for allocation, if not specific, sum by svr_N_i
  + @param fl_rho float preference for equality for the planner
  + @return a dataframe that expands the df inputs with additional results.
2. The structure assumes some regression has already taken place to generate the i specific variables listed. and

Doing this allows for lagged intereaction that are time specific in an arbitrary way.

## Set Up

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(dplyr)
library(tidyr)
library(stringr)
library(broom)
library(ggplot2)
library(R4Econ)
library(knitr)
library(kableExtra)
```

## Get Data

Generate four categories by initial height and mother's education levels combinations.

```{r Load Packages and Process Data}
# Load Library

# Select Cebu Only
df_hw_cebu_m24 <- df_hgt_wgt %>% filter(S.country == 'Cebu' & svymthRound == 24 & prot > 0 & hgt > 0) %>% drop_na()

# Generate Discrete Version of momEdu
df_hw_cebu_m24 <- df_hw_cebu_m24 %>%
    mutate(momEduRound = cut(momEdu,
                             breaks=c(-Inf, 10, Inf),
                             labels=c("MEduLow","MEduHigh"))) %>%
    mutate(hgt0med = cut(hgt0,
                             breaks=c(-Inf, 50, Inf),
                             labels=c("h0low","h0high")))

df_hw_cebu_m24$momEduRound = as.factor(df_hw_cebu_m24$momEduRound)
df_hw_cebu_m24$hgt0med = as.factor(df_hw_cebu_m24$hgt0med)

# Attach
attach(df_hw_cebu_m24)
```

# Regression with Data and Construct Input Arrays

## Linear Regression

Estimation *Production functions* or any other function. Below, Regression height outcomes on input interacted with the categories created before. This will generate category specific marginal effects.

```{r Linear Regression}
# Input Matrix
mt_lincv <- model.matrix(~ hgt0 + wgt0)
mt_linht <- model.matrix(~ sex:hgt0med - 1)

# Regress Height At Month 24 on Nutritional Inputs with controls
rs_hgt_prot_lin = lm(hgt ~ prot:mt_linht + mt_lincv - 1)
print(summary(rs_hgt_prot_lin))
rs_hgt_prot_lin_tidy = tidy(rs_hgt_prot_lin)
```

## Log-Linear Regression

Now log-linear regressions, where inptut coefficient differs by input groups.

```{r Log-Linear Regression}
# Input Matrix Generation
mt_logcv <- model.matrix(~ hgt0 + wgt0)
mt_loght <- model.matrix(~ sex:hgt0med - 1)

# Log and log regression for month 24
rs_hgt_prot_log = lm(log(hgt) ~ log(prot):mt_loght + mt_logcv - 1)
print(summary(rs_hgt_prot_log))
rs_hgt_prot_log_tidy = tidy(rs_hgt_prot_log)
```

## Construct Input Arrays $A_i$ and $\alpha_i$

Multiply coefficient vector by covariate matrix to generate *A* vector that is child/individual specific.

```{r Post Regression Input Processing}

# Generate A_i
ar_Ai_lin <- mt_lincv %*% as.matrix(rs_hgt_prot_lin_tidy %>% filter(!str_detect(term, 'prot')) %>% select(estimate))
ar_Ai_log <- mt_logcv %*% as.matrix(rs_hgt_prot_log_tidy %>% filter(!str_detect(term, 'prot')) %>% select(estimate))

# Generate alpha_i
ar_alphai_lin <- mt_linht %*% as.matrix(rs_hgt_prot_lin_tidy %>% filter(str_detect(term, 'prot')) %>% select(estimate))
ar_alphai_log <- mt_loght %*% as.matrix(rs_hgt_prot_log_tidy %>% filter(str_detect(term, 'prot')) %>% select(estimate))

# Child Weight
ar_beta <- rep(1/length(ar_Ai_lin), times=length(ar_Ai_lin))

# Initate Dataframe that will store all estimates and optimal allocation relevant information
mt_opti <- cbind(ar_alphai_lin, ar_Ai_lin, ar_beta)
ar_st_varnames <- c('alpha', 'A', 'beta')
tb_opti <- as_tibble(mt_opti) %>% rename_all(~c(ar_st_varnames))

# Unique beta, A, and alpha groups
tb_opti_unique <- tb_opti %>% group_by(!!!syms(ar_st_varnames)) %>%
                    arrange(!!!syms(ar_st_varnames)) %>%
                    summarise(n_obs_group=n())
```

# Optimal Allocations

## Common Parameters for Optimal Allocation

```{r Set Allocation Parameters}
# Child Count
df_hw_cebu_m24_full <- df_hw_cebu_m24
it_obs = dim(df_hw_cebu_m24)[1]

# Total Resource Count
ar_prot_data = df_hw_cebu_m24$prot
fl_N_agg = sum(ar_prot_data)

# Vector of Planner Preference
ar_rho = c(seq(-200, -100, length.out=5), seq(-100, -25, length.out=5), seq(-25, -5, length.out=5), seq(-5, -1, length.out=5), seq(-1, -0.01, length.out=5), seq(0.01, 0.25, length.out=5), seq(0.25, 0.99, length.out=5))
ar_rho = c(-50, -25, -10)
ar_rho = unique(ar_rho)
```

## Optimal Linear Allocation (CRS)

This also works with any CRS CES.

### Optimal Linear Allocation Hard-Coded

```{r Optimal Linear Allocation Hard Code All Rho}
# Optimal Linear Equation
# Planner Inputs
mt_hev_lin = matrix(, nrow = length(ar_rho), ncol = 2)
mt_opti_N = matrix(, nrow = it_obs, ncol = length(ar_rho))

# A. First Loop over Planner Preference
# Generate Rank Order
for (it_rho_ctr in seq(1,length(ar_rho))) {
  rho = ar_rho[it_rho_ctr]

  # B. Generate V4, Rank Index Value, rho specific
  # tb_opti <- tb_opti %>% mutate(!!paste0('rv_', it_rho_ctr) := A/((alpha*beta))^(1/(1-rho)))
  tb_opti <- tb_opti %>% mutate(rank_val = A/((alpha*beta))^(1/(1-rho)))

  # c. Generate Rank Index
  tb_opti <- tb_opti %>% arrange(rank_val) %>% mutate(rank_idx = row_number())

  # d. Populate lowest index alpha, beta, and A to all rows
  tb_opti <- tb_opti %>% mutate(lowest_rank_A = A[rank_idx==1]) %>%
                mutate(lowest_rank_alpha = alpha[rank_idx==1]) %>%
                mutate(lowest_rank_beta = beta[rank_idx==1])

  # e. relative slope and relative intercept with respect to lowest index
  tb_opti <- tb_opti %>%
                mutate(rela_slope_to_lowest =
                         (((lowest_rank_alpha*lowest_rank_beta)/(alpha*beta))^(1/(rho-1))*(lowest_rank_alpha/alpha))
                      ) %>%
                mutate(rela_intercept_to_lowest =
                         ((((lowest_rank_alpha*lowest_rank_beta)/(alpha*beta))^(1/(rho-1))*(lowest_rank_A/alpha)) - (A/alpha))
                      )

  # f. cumulative sums
  tb_opti <- tb_opti %>%
                mutate(rela_slope_to_lowest_cumsum =
                         cumsum(rela_slope_to_lowest)
                      ) %>%
                mutate(rela_intercept_to_lowest_cumsum =
                         cumsum(rela_intercept_to_lowest)
                      )

  # g. inverting cumulative slopes and intercepts
  tb_opti <- tb_opti %>%
                mutate(rela_slope_to_lowest_cumsum_invert =
                         (1/rela_slope_to_lowest_cumsum)
                      ) %>%
                mutate(rela_intercept_to_lowest_cumsum_invert =
                         ((-1)*(rela_intercept_to_lowest_cumsum)/(rela_slope_to_lowest_cumsum))
                      )

  # h. Relative x-intercept points
  tb_opti <- tb_opti %>%
                mutate(rela_x_intercept =
                         (-1)*(rela_intercept_to_lowest/rela_slope_to_lowest)
                      )

  # i. Inverted relative x-intercepts
  tb_opti <- tb_opti %>%
                mutate(opti_lowest_spline_knots =
                         (rela_intercept_to_lowest_cumsum + rela_slope_to_lowest_cumsum*rela_x_intercept)
                      )

  # j. Sort by order of receiving transfers/subsidies
  tb_opti <- tb_opti %>% arrange(rela_x_intercept)

  # k. Find position of subsidy
  tb_opti <- tb_opti %>% arrange(opti_lowest_spline_knots) %>%
                mutate(tot_devi = opti_lowest_spline_knots - fl_N_agg) %>%
                arrange((-1)*case_when(tot_devi < 0 ~ tot_devi)) %>%
                mutate(allocate_lowest =
                         case_when(row_number() == 1 ~
                                     rela_intercept_to_lowest_cumsum_invert +
                                     rela_slope_to_lowest_cumsum_invert*fl_N_agg)) %>%
                mutate(allocate_lowest = allocate_lowest[row_number() == 1]) %>%
                mutate(opti_allocate =
                         rela_intercept_to_lowest +
                         rela_slope_to_lowest*allocate_lowest) %>%
                mutate(opti_allocate =
                         case_when(opti_allocate >= 0 ~ opti_allocate)) %>%
                mutate(opti_allocate_total = sum(opti_allocate, na.rm=TRUE))
  
  # l. Summarize outputs
  tb_opti_main_results <- tb_opti %>% 
    select(-one_of(c('lowest_rank_alpha', 'lowest_rank_beta')))
  
  
  print('')
  print('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')
  print(paste0('xxx rho:', rho))
  print('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')
  print('xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx')
  
  print(summary(tb_opti_main_results))

}

# lineplot <- tb_opti %>%
#     gather(variable, value, -month) %>%
#     ggplot(aes(x=month, y=value, colour=variable, linetype=variable)) +
#         geom_line() +
#         geom_point() +
#         labs(title = 'Mean and SD of Temperature Acorss US Cities',
#              x = 'Months',
#              y = 'Temperature in Fahrenheit',
#              caption = 'Temperature data 2017') +
#         scale_x_continuous(labels = as.character(df_temp_mth_summ$month),
#                            breaks = df_temp_mth_summ$month)

```
